### 2014-08-07 disregarding-ngrams.Rmd
Measure quality of OCR

---
title: "How many n-grams are we disregarding?"
output: html_document
---

```{r, echo=FALSE}
library(magrittr)
library(RWeka)
library(stringr)
```

```{r, echo=FALSE}
codes_dir <- "text"
files <- dir(codes_dir, "*.txt")
raw <- file.path(codes_dir, files) %>%
  lapply(., scan, "character", sep = "\n")
names(raw) <- files
codes_texts <- lapply(raw, paste, collapse = " ") %>%
  lapply(., tolower) %>%
  lapply(., WordTokenizer) %>%
  lapply(., paste, collapse = " ")
```

```{r, echo=FALSE}
ngrammify <- function(data, n) { 
  NGramTokenizer(data, Weka_control(min = n, max = n))
  }
```

```{r, echo=FALSE, cache=TRUE}
codes_grams <- lapply(codes_texts, ngrammify, 5)
```

```{r, echo=FALSE}
#' Remove unreasonable n-grams containing characters other than letters and spaces
#' @param ngrams A list of n-grams
#' @return Returns a list of filtered n-grams
filter_unreasonable_ngrams <- function(ngrams) {
  require(stringr)
  ngrams[!str_detect(ngrams, "[^a-z ]")]
}
```

We have a known problem of mistakes in the OCR coupled with section numbers, page numbers, and the occasional stray character. If we throw away any n-gram that includes a character outside the range of `a-z` (plus space), what proportion of the n-grams is left? This is a rough proxy for how good the OCR is.

```{r}
codes_grams %>% lapply(function(x) {
  (x %>% filter_unreasonable_ngrams %>% length) / (x %>% length)
}) 
```

Using tessaract instead of other OCR solutions, we're keeping about 80% of the NY 1850 code, instead of roughly 70% using our older method. Since that number is the denominator when we calculate the proportion of the code that has a match, this explains why the proportion grew less than we might have expected.