### 2014-08-15 test-compare-codes.Rmd
Use unique n-grams for the denominator
Since `intersect()` finds the unique n-grams that match, not all the
matches, it is necessary to use the unique n-grams in the denominator
too. This shows that the comparisons are more favorable than we thought.
And most important, in the diagonal of the comparison matrix, a code
compared to itself now has a value of 1.

---
title: "First Attempt at a Comparison Matrix"
output: html_document
---

Let's do a quick and dirty test to see if we can get a comparison matrix.

```{r}
library(magrittr)
library(RWeka)
library(stringr)
```

Load the codes.

```{r, echo=FALSE}
codes_dir <- "text"
files <- dir(codes_dir, "*.txt")
raw <- file.path(codes_dir, files) %>%
  lapply(., scan, "character", sep = "\n")
names(raw) <- files
codes_texts <- lapply(raw, paste, collapse = " ") %>%
  lapply(., tolower) %>%
  lapply(., WordTokenizer) %>%
  lapply(., paste, collapse = " ")

```

Keep a random sample of three plus NY 1850 to keep computation time low for the test.

```{r}
codes_sample <- c(sample(codes_texts, 3), NY1850.txt = codes_texts[["NY1850.txt"]])
```

Function for n-grams:

```{r}
ngrammify <- function(data, n) { 
  NGramTokenizer(data, Weka_control(min = n, max = n))
  }
```

Create the n-grams:

```{r}
codes_grams <- lapply(codes_sample, ngrammify, 5)
```

Function to remove unreasonable n-grams:

```{r}
#' Remove unreasonable n-grams containing characters other than letters and spaces
#' @param ngrams A list of n-grams
#' @return Returns a list of filtered n-grams
filter_unreasonable_ngrams <- function(ngrams) {
  require(stringr)
  ngrams[!str_detect(ngrams, "[^a-z ]")]
}
```

Now apply that function to `codes_grams` so we get a list of clean n-grams. This might not be the best strategy: we'll think about that later.

```{r}
codes_grams <- lapply(codes_grams, filter_unreasonable_ngrams)
```

Function to compare two codes. Note that this function (unlike earlier versions and just for the purpose of this test) only returns one number for simplicity's sake instead of a list. (And it no longer applies the filter for unreasonable n-grams inside this function. Instead that is applied globally earlier.)

```{r}
#' Compare two codes using their list of n-grams.
#' @param orig_code A vector of n-grams representing a code.
#' @param dest_code A vectory of n-grams representing a code.
#' @return A list containing data that might help to identify whether codes
#'   match up.
compare_codes_by_shared_ngrams <- function(orig_code, dest_code) {
  require(magrittr)

  matches <- intersect(orig_code, dest_code)
  shared_ngrams <- unique(c(orig_code, dest_code))
  ratio_matches_to_possible <- length(matches) / length(unique(shared_ngrams))
  ratio_matches_to_destination <- length(matches) / length(unique(dest_code))
  return(ratio_matches_to_destination)
}
```

Now we use a function to create a matrix of comparisons (This is inefficient because it compares each code to itself. And for each possible pairing it does the comparison twice, swapping origin and destination codes. Good enough for now.)

This is to help us understand the comparisons (pretty sure I have the arrow pointing the right way!):

```{r}
code_names <- str_extract(names(codes_grams), "\\w+")
names_of_comparisons <- outer(code_names, code_names, function(x, y){
  paste(y, "from", x)
}) 
colnames(names_of_comparisons) <- code_names
rownames(names_of_comparisons) <- code_names
print(names_of_comparisons)
```

These are the comparisons themselves. The way to read this matrix is "the proportion of ngrams that COLUMN borrows from ROW." In other words the ROW is the origin code and the COLUMN is the destination code.

```{r}
# From http://stackoverflow.com/questions/1719447/outer-equivalent-for-non-vector-lists-in-r
outer_for_lists <- function(a,b, fun) {
  outer(a, b, function(x,y) vapply(seq_along(x), function(i) fun(x[[i]], y[[i]]), numeric(1)))
}
comparison_of_codes <- outer_for_lists(codes_grams, codes_grams, compare_codes_by_shared_ngrams)
colnames(comparison_of_codes) <- code_names
rownames(comparison_of_codes) <- code_names
print(comparison_of_codes)
```

Now we can run this analysis on the whole shebang. First we get all the codes as n-grams:

```{r}
codes_all_grams <- lapply(codes_texts, ngrammify, 5)
codes_all_grams <- lapply(codes_all_grams, filter_unreasonable_ngrams)
```

Here is the matrix of comparisons for all the codes:

```{r}
codes_all_names <- str_extract(names(codes_all_grams), "\\w+")
comparison_of_all_codes <- outer_for_lists(codes_all_grams, codes_all_grams, compare_codes_by_shared_ngrams)
colnames(comparison_of_all_codes) <- codes_all_names
rownames(comparison_of_all_codes) <- codes_all_names
print(comparison_of_all_codes)
```

Let's write that comparison to disk for someone who know about the history of codes to make sense of it.

```{r}
write.csv(comparison_of_all_codes, "out/comparison_of_all_codes.csv")
```

What we really care about is just the triangular comparison.

```{r}
triangle <- comparison_of_all_codes
triangle[upper.tri(triangle, diag = TRUE)]  <- NA
print(triangle)
```

What is the distribution of comparisons?

```{r}
triangle %>% hist()
```

For each code we can get a sorted list of the codes that it is most like.

```{r}
for(i in 1:ncol(comparison_of_all_codes)) {
  x <- comparison_of_all_codes[,i]
  title <- colnames(comparison_of_all_codes)[i]
  cat("\n\n", title, " is most like:\n", sep = "")
  cat("---------\n")
  y <- sort(x, decreasing = TRUE)
  for(j in 2:length(y)) {
    cat(names(y)[j], y[j], "\n", sep = " ") 
    }
}
```

*********

Now this is more than we can figure out meaningfully today, but it's possible to run standard distance and clustering functions on that matrix. I'm not sure what this even means just yet.

Let's see the standard distance function:

```{r}
distance <- comparison_of_all_codes %>% dist
print(distance)
```

And we can then pass that distance matrix to hierarchical clustering or k-means clustering (obviously we need to read up on all these) and plot each.

```{r}
distance %>% hclust() %>% str()
distance %>% hclust() %>% plot()
```

I wouldn't put any stock in that chart, but it does show what should be possible. And apart from the problem of the chronological arrow, it is interesting that western states are close to one another, the New York codes are close together, the two Ohio codes are at least close together, etc. Time to learn some statistics.

And now for k-means with a plot that I kind of get.

```{r}
cluster <- kmeans(comparison_of_all_codes, centers = 5)
library(cluster)
clusplot(comparison_of_all_codes, cluster$cluster, labels = 2)
```