### 2014-11-21 network-analysis.Rmd
Remove anachronistic comparisons

---
title: "Network Analysis of Code Similarity"
author: "Kellen Funk and Lincoln Mullen"
date: "November 21, 2014"
output: html_document
---

```{r echo=FALSE}
packrat::on()
library(magrittr)
library(RWeka)
library(stringi)
source("R/helper.R")

# Load the data if it doesn't exist already.
if(file.exists("out/codes.rda")) {
  load("out/codes.rda")
  } else {
  source("R/load-data.R")
  save(codes, file = "out/codes.rda")
  }
```

There are the codes that we currently have:

```{r}
names(codes)
```

Create the comparison matrix:

```{r}
cf <- outer_for_lists(codes, codes, compare_codes_by_shared_ngrams)
diag(cf) <- NA                 # remove codes compared to themselves
cf <- rm_anachronism(cf)       # remove anachronistic comparisons
```