### 2014-11-21 network-analysis.Rmd
Basic network graph

---
title: "Network Analysis of Code Similarity"
author: "Kellen Funk and Lincoln Mullen"
date: "November 21, 2014"
output: html_document
---

```{r echo=FALSE}
packrat::on()
library(magrittr)
library(RWeka)
library(stringi)
library(igraph)
source("R/helper.R")

# Load the data if it doesn't exist already.
if(file.exists("out/codes.rda")) {
  load("out/codes.rda")
  } else {
  source("R/load-data.R")
  save(codes, file = "out/codes.rda")
  }
```

There are the codes that we currently have:

```{r}
names(codes)
```

Create the comparison matrix:

```{r}
cf <- outer_for_lists(codes, codes, compare_codes_by_shared_ngrams)
```

Remove obviously incorrect values:

```{r}
cf2  <- cf
diag(cf2) <- NA                 # remove codes compared to themselves
cf2 <- rm_anachronism(cf2)       # remove anachronistic comparisons

significance_threshold <- 0.15
cf2[cf2 < significance_threshold] <- NA
cf2[is.na(cf2)] <- 0
```

Now we can create a graph object from the comparison matrix, which we will treat as an adjacency matrix.

```{r}
g <- graph.adjacency(cf2, mode = "directed", weighted = TRUE, diag = FALSE) 
set.seed(4821)
l <- layout.fruchterman.reingold(g, weights = E(g)$weight)
par(mar=c(0,0,0,0))
plot(g, edge.width = E(g)$weight * 12, layout = l, edge.arrow.size= 0.25, vertex.size = 5)
```