### 2015-09-21 early-analysis/006-ocr-quality.Rmd
Move early analyses into separate folder

---
title: "Check OCR quality of specific codes"
author: "Lincoln Mullen"
date: "September 11, 2015"
output: html_document
---

We have multiple versions of several codes, so we want to see whether any of them have better quality than others. The codes we are interested in:

> For Iowa 1859, I have a, b, and c versions, and for New Mexico 1897 I have an a and b.

```{r, message=FALSE, warning=FALSE}
library(ocrquality)
library(stringr)
library(dplyr)
set.seed(33532)
```

Load the files.

```{r}
paths <- Sys.glob("legal-codes/*.txt")
docs <- paths %>% 
  lapply(readLines) %>% 
  lapply(str_c, collapse = " ")
```

Check the OCR quality:

```{r}
quality <- sapply(docs, ocr_dictionary, sample_size = 10e3)
```

Turn that into a data frame and print it.

```{r}
results <- data_frame(file = paths,
                      quality = quality) %>% 
  arrange(desc(quality))
relevant <- results %>% 
  filter(str_detect(file, "IA1859") | str_detect(file, "NM1897"))
relevant
```

Write to CSV.

```
write.csv(results, "out/ocr-quality.csv")
write.csv(relevant, "out/ocr-quality-relevant.csv")
```