### 2015-09-21 007-load-corpus.Rmd
Load and tokenize the corpus

---
title: "Load and tokenize all codes"
author: "Lincoln Mullen"
date: "September 21, 2015"
output: html_document
---

We are going to tokenize the codes and save a cached version. For our initial analysis, we will use skip n-grams with a moderate value for `k` so that we have many features.

```{r}
library("textreuse")

cache <- "cache/corpus-skip-ngrams-n7k3.rds"
if (!file.exists(cache)) {
  description <- "General purpose tokenization of codes for comparing pairwise."
  corpus <- TextReuseCorpus(dir = "legal-codes/",
                            meta = list("description" = description,
                                        "title" = "Field codes"),
                            tokenizer = tokenize_skip_ngrams, n = 7 , k = 3,
                            hash_func = hash_string,
                            keep_tokens = TRUE)
  dir.create("cache", showWarnings = FALSE)
  corpus
  saveRDS(corpus, file = cache)
}
```

Session info:

```{r}
sessionInfo()
```