### 2015-09-21 008-pairwise-jaccard.Rmd
Pairwise with Jaccard similarity

---
title: "Compare pairwise with Jaccard similarity"
author: "Lincoln Mullen"
date: "September 21, 2015"
output: 
  html_document:
    keep_md: true
---

We are going to load the corpus from the cache, and create a comparison matrix with Jaccard similarity. (We will do other comparisons in different notebook so we can do this in parallel.) We will cache the matrix for use in later analyses.

```{r}
library("textreuse")

cache <- "cache/comparison-matrix-jaccard.rds"
if (!file.exists(cache)) {
  corpus <- readRDS("cache/corpus-skip-ngrams-n7k3.rds")
  m <- pairwise_compare(corpus, jaccard_similarity)
  saveRDS(m, cache)
} else {
  m <- readRDS(cache)
}
```

Here is the comparison matrix:

```{r}
m
```

Session info:

```{r}
sessionInfo()
```