### 2015-09-23 011-sections-ny-ca.Rmd
Testing LSH on sectioned codes

---
title: "LSH for sections of CA and NY"
author: "Lincoln Mullen"
date: "September 22, 2015"
output: 
  html_document:
    keep_md: true
---

```{r, message=FALSE}
library("textreuse")
library("stringr")
library("dplyr")
source("R/extract_code_names.R")
```

As a test run for using LSH on all the codes, we are going to run this on just NY1850 and CA1851 and see if we can detect matches between sections of that code.

We begin by identifying all of the sections of those two codes:

```{r}
paths <- c(Sys.glob("legal-codes-split/NY1850*"), 
           Sys.glob("legal-codes-split/CA1851*"))
```

That is `r prettyNum(length(paths), big.mark = ",")` sections. We want to eliminate extremely small sections. So we are going to check the sizes of all of the files. Since we know that a file with five words has 32 bytes (using `du -b`) we will remove all files with fewer than 32 bytes. This will also take care of the empty `-SPLIT` files that we use for Make tasks. (Addendum, in the future we will do this by checking the word count directly.)

```{r}
sizes <- file.size(paths)
paths <- paths[sizes >= 34]
```

Now we have `r prettyNum(length(paths), big.mark = ",")` sections.

## Loading, tokenizing, and minhashing

We need an appropriate threshold for LSH. We are aiming for greater than `0.1`. So these values seem give us a slightly higher threshold.

```{r}
lsh_threshold(1500, 500)
```

We load and tokenize those files with 4-grams.

```{r}
minhash <- minhash_generator(1500, seed = 343653)
corpus <- TextReuseCorpus(paths = paths, 
                          tokenizer = tokenize_ngrams, n = 4,
                          hash_func = minhash,
                          keep_tokens = TRUE)
```

Now we can run LSH on the corpus to find the matches.

```{r}
buckets <- lsh(corpus, bands = 500)
candidates <- lsh_candidates(buckets) 
corpus <- rehash(corpus, hash_string)
scores <- lsh_compare(candidates, corpus, jaccard_similarity)
scores_filtered <- scores %>% 
  mutate(code_a = extract_code_names(a),
         code_b = extract_code_names(b)) %>% 
  filter(code_a != code_b)
```

We have `r nrow(scores_filtered)` matches between the two codes. If we filter so that we only keep track of matches with a similarity of `0.1`, there are `r nrow(filter(scores_filtered, score >= 0.1))`. For comparison, there are 691 sections in the split CA1851 code.

## Conclusions

We have genuine matches, even on the low Jaccard scores. 4-grams are too short because phrases like "in the same manner" and "in a civil action" are very common. We think 5-grams are the best length to avoid common phrases. The best threshold for distinguishing genuine matches from false positives is `0.1`.