### 2015-09-25 017-best-section-matches.Rmd
Investigate section matches by code

---
title: "Best matches for code sections"
author: "Lincoln Mullen"
date: "September 25, 2015"
output: 
  html_document:
    keep_md: true
---

```{r message=FALSE}
library("dplyr")
library("stringr")
library("readr")
library("ggplot2")
library("RColorBrewer")
source("R/extract-date.R")
source("R/extract_code_names.R")
```

The aim of this notebook is to create a function that takes a code's abbreviation (e.g., `NY1850`) and creates a data frame with each of the sections in that code with the best match to a non-anachronistic code.

Read in the data.

```{r}
scores <- read_csv("out/scores-all-sections-pairs.csv")
scores <- scores %>% 
  rename(section_a = a, section_b = b) %>% 
  mutate(code_a = extract_code_names(section_a),
         code_b = extract_code_names(section_b),
         year_a = extract_date(code_a),
         year_b = extract_date(code_b))
scores
```

```{r}
best_section_matches <- function(code_name, scores, threshold = 0.1, top = 1) {
  require("dplyr")
  require("stringr")
  matches <- scores %>% 
    filter(code_a == code_name | code_b == code_name) %>% 
    mutate(match_code = ifelse(code_a == code_name, code_b, code_a),
           code_of_interest = ifelse(code_a == code_name, code_a, code_b),
           match_section = ifelse(code_a == code_name, section_b, section_a),
           section_of_interest = ifelse(code_a == code_name, section_a, section_b),
           match_year = ifelse(code_a == code_name, year_b, year_a)) %>% 
    filter(match_code != code_of_interest,
           match_year <= extract_date(code_name),
           similarity >= threshold) %>% 
    select(code_of_interest, section_of_interest, match_code, match_section,
           similarity, dissimilarity) %>% 
    group_by(section_of_interest) %>%
    arrange(desc(similarity)) %>% 
    top_n(top, similarity)

  all_sections <- Sys.glob(str_c("legal-codes-split/", code_name, "-0*")) %>% 
    str_replace("legal-codes-split/", "") %>% 
    str_replace("\\.txt", "")
  all <- data_frame(all_sections)

  all %>% 
    left_join(matches, by = c("all_sections" = "section_of_interest")) %>% 
    select(-code_of_interest)
} 
```

Also a function to calculate summary statistics.

```{r}
summarize_borrowings <- function(section_list) {
  section_list %>% 
    group_by(match_code) %>% 
    summarize(mean_similarity = mean(similarity),
              n = n()) %>% 
    mutate(percentage_sections = n / nrow(section_list)) %>% 
    arrange(desc(n))
}
```

Apply this to three codes:

```{r}
CA1851 <- best_section_matches("CA1851", scores, threshold = 0.2)
CA1851
summarize_borrowings(CA1851)
IA1859 <- best_section_matches("IA1859", scores, threshold = 0.2)
IA1859
summarize_borrowings(IA1859)
UT1870 <- best_section_matches("UT1870", scores, threshold = 0.2)
UT1870
summarize_borrowings(UT1870)
```

Make a plot of code borrowings.

```{r}
plot_borrowings <- function(state_borrowings) {
  state_borrowings %>% 
    mutate(state = str_extract(match_code, "\\w\\w")) %>% 
    ggplot(aes(x = as.factor(all_sections), y = similarity, fill = state)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_x_discrete(labels = NULL) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_brewer(palette = "Dark2")
}
```

Plot some borrowings:

```{r}
plot_borrowings(CA1851)
plot_borrowings(UT1870)
```