### 2015-09-26 018-families-of-sections.Rmd
Families of sections

---
title: "Clustering"
author: "Lincoln Mullen"
date: "September 25, 2015"
output: 
  html_document:
    keep_md: true
---

```{r, message = FALSE}
library("readr")
library("dplyr")
library("igraph")
```

How are all of the sections related to one another? We can think of this as a network problem. Sections that are similar have a relationship, but are very unlikely to have relationships to every other section. So we can use our comparisons then decompose that big network graph into families of codes.

Read the data.

```{r}
comparisons <- read_csv("out/scores-all-sections-pairs.csv") %>% 
  select(-dissimilarity) %>% 
  rename(weight = similarity) %>% 
  filter(weight > 0.1)
```

```{r}
g <- graph.data.frame(comparisons, directed = "FALSE")
families <- decompose(g, min.vertices = 2)
num_sections <- vapply(families, function(x) { length(V(x))}, integer(1))
head(sort(-num_sections), 20)
which(num_sections > 10 & num_sections < 15)
```

Plot those and look at their vertices.

```{r}
V(families[[4]])
plot(families[[4]])
V(families[[1084]])
plot(families[[1084]])
V(families[[4635]])
plot(families[[4635]])
```