### 2015-10-08 008-pairwise-jaccard.Rmd
Simplify notebook generation, improve figures

---
title: "Compare pairwise with Jaccard similarity"
author: "Lincoln Mullen"
date: "September 21, 2015"
---

```{r}
library("textreuse")
```

We are going to load the corpus from the cache, and create a comparison matrix with Jaccard similarity. (We will do other comparisons in different notebook so we can do this in parallel.) We will cache the matrix for use in later analyses.

We are going to test this with both shingled and skip n-grams. Here are skip n-grams.

```{r}
cache_skip <- "cache/comparison-matrix-skip-jaccard.rds"
if (!file.exists(cache_skip)) {
  corpus_skip <- readRDS("cache/corpus-skip-ngrams-n7k3.rds")
  m_skip <- pairwise_compare(corpus_skip, jaccard_similarity)
  saveRDS(m_skip, cache_skip)
} else {
  m_skip <- readRDS(cache_skip)
}
```

Here are shingled n-grams.

```{r}
cache_shingles <- "cache/comparison-matrix-shingles-jaccard.rds"
if (!file.exists(cache_shingles)) {
  corpus_shingles <- readRDS("cache/corpus-shingle-ngrams-n7.rds")
  m_shingles <- pairwise_compare(corpus_shingles, jaccard_similarity)
  saveRDS(m_shingles, cache_shingles)
} else {
  m_shingles <- readRDS(cache_shingles)
}
```

Here is the comparison matrix for the skip n-grams.

```{r}
round(m_skip, digits = 4)
```

Here is the comparison matrix for the shingled n-grams.

```{r}
round(m_shingles, digits = 4)
```

Session info:

```{r}
sessionInfo()
```