### 2015-10-08 016-manual-lsh.Rmd

---
title: "Calculate minhash pairs manually"
author: "Lincoln Mullen"
date: "September 25, 2015"
---

As the textreuse package is currently implemented, the `lsh()` function gets bogged down with large numbers of documents. The implementation with the hash package is probably unneccessary; we can do it a lot faster with dplyr. So we will implement a specific solution here before revising the textreuse package. (Thanks for an assist from [Stack Overflow](http://stackoverflow.com/questions/32783727/turn-ordered-pairs-into-unordered-pairs-in-a-data-frame-with-dplyr/32784685#32784685).)

```{r message=FALSE}
library("readr")
library("textreuse")
library("dplyr")
library("digest")
library("stringr")
library("tidyr")
```

Load the corpus.

```{r}
cache <- "cache/corpus-n5-minhash-h1500-seed623.rds"
if (!file.exists(cache)) {
  corpus <- TextReuseCorpus(dir = "legal-codes-split", tokenizer = NULL)
  wc <- wordcount(corpus)
  corpus <- corpus[wc > 5]
  minhash <- minhash_generator(n = 1500, seed = 623)
  corpus <- tokenize(corpus, tokenizer = tokenize_ngrams, n = 5,
                     hash_func = minhash, keep_tokens = TRUE)
  saveRDS(corpus, file = cache)
} else {
  corpus <- readRDS(cache)
}
```

```{r}
bands <- 500
hashes <- 1500
rows <- hashes / bands
docs <- length(corpus)
bands_vec <- rep(vapply(1:bands, function(i) rep(i, rows), integer(rows)), docs)

buckets_cache <- "cache/buckets-dplyr-lsh.rds"
if (!file.exists(buckets_cache)) {
  buckets <-  corpus %>%
    hashes() %>%
    as_data_frame() %>%
    gather(doc, hash) %>%
    mutate(doc = as.character(doc),
           band = bands_vec) %>%
    group_by(doc, band) %>%
    summarize(buckets = digest(hash)) %>%
    select(-band)
  rm(bands_vec)
  saveRDS(buckets, file = buckets_cache)
} else {
  buckets <- readRDS(buckets_cache)
}
```

Now find the ordered pairs of matches:

```{r}
ordered <- buckets %>% 
  ungroup() %>%
  left_join(buckets, by = "buckets") %>%
  select(-buckets) %>%
  distinct() %>% 
  filter(doc.x != doc.y) %>%
  arrange(doc.x, doc.y)
```

Now remove the duplicate pairs.

```{r}
unordered <- ordered %>% 
  distinct(dn = pmin(doc.x, doc.y), up = pmax(doc.x, doc.y)) %>%
  select(-up, -dn) %>%
  rename(a = doc.x, b = doc.y) %>%
  arrange(a, b) %>% 
  mutate(score = NA_real_) 
```

Save the results.

```{r}
write_csv(ordered, path = "out/all-sections-unordered-pairs.csv")
```

Rehash the corpus to use for Jaccard similarity.

```{r}
corpus_rehashed <- rehash(corpus, hash_string)
```

Compute the Jaccard similarity scores and save the results.

```{r}
scores <- unordered %>% 
  lsh_compare(corpus_rehashed, jaccard_similarity) %>% 
  rename(similarity = score) %>% 
  mutate(dissimilarity = 1 - similarity)
write_csv(scores, path = "out/scores-all-sections-pairs.csv")
scores
```