### 2015-10-30 020-lsh-with-package.Rmd
LSH the corpus

---
title: "Use package"
author: "Lincoln Mullen"
date: "October 19, 2015"
output: html_document
---

We have refined our technique for extracting matches. So the work in the document supercedes all previous work. The aim is to create the data frame of matches and other results and cache them for future use.

```{r}
library("textreuse")
library("dplyr")
options("mc.cores" = 6L)
```

Create the corpus.

```{r}
cache_corpus <- "cache/corpus-sections-minash-n1500-seed623.rda"
if (!file.exists(cache_corpus)) {
  message("Creating the corpus")
  minhash <- minhash_generator(n = 1500, seed = 623)
  sections <- TextReuseCorpus(dir = "legal-codes-split/",
                              tokenizer = tokenize_ngrams, n = 5,
                              keep_tokens = FALSE,
                              minhash_func = minhash)
  save(sections, minhash, file = cache_corpus)
} else {
  message("Loading corpus from cache")
  load(file = cache_corpus)
}
```

Compute the LSH candidates, then compare them.

```{r}
cache_lsh <- "cache/lsh-sections.rda"
if (!file.exists(cache_lsh)) {
  message("Running LSH analysis")
  section_buckets <- lsh(sections, bands = 500)
  section_candidates <- lsh_candidates(section_buckets)
  section_scores <- lsh_compare(section_candidates, sections, jaccard_similarity)
  save(section_buckets, section_candidates, section_scores, file = cache_lsh)
  saveRDS(section_scores, "cache/section-scores.rds")
} else {
  message("Loading LSH objects from cache")
  load(file = cache_lsh)
}
```

How many pairs of matches do we have above `0.1` Jaccard score?

```{r}
section_scores %>% 
  filter(score > 0.1) %>% 
  nrow()
```