### 2015-10-31 020-lsh-with-package.Rmd
Redo the LSH with new sections and smarter minhashes/bands

---
title: "Use package"
author: "Lincoln Mullen"
date: "October 19, 2015"
output: html_document
---

We have refined our technique for extracting matches. So the work in the document supercedes all previous work. The aim is to create the data frame of matches and other results and cache them for future use.

```{r message=FALSE}
library("textreuse")
library("dplyr")
options("mc.cores" = 6L)
```

We want to have the smallest number of hashes and bands while still detecting all reasonable matchs. We will set the bands and number of minhashes as follows.

```{r}
h <- 120
b <- 60
lsh_threshold(h, b)
```

Create the corpus.

```{r}
cache_corpus <- "cache/corpus-sections-minash-n120-seed623.rda"
if (!file.exists(cache_corpus)) {
  message("Creating the corpus.")
  minhash <- minhash_generator(n = h, seed = 623)
  sections <- TextReuseCorpus(dir = "legal-codes-split/",
                              tokenizer = tokenize_ngrams, n = 5,
                              keep_tokens = FALSE,
                              minhash_func = minhash)
  save(sections, minhash, file = cache_corpus)
} else {
  message("Loading corpus from cache.")
  load(file = cache_corpus)
}
```

Compute the LSH candidates, then compare them.

```{r}
cache_lsh <- "cache/lsh-sections.rda"
if (!file.exists(cache_lsh)) {
  message("Running LSH analysis.")
  section_buckets <- lsh(sections, bands = b)
  section_candidates <- lsh_candidates(section_buckets)
  section_scores <- lsh_compare(section_candidates, sections, jaccard_similarity)
  save(section_buckets, section_candidates, section_scores, file = cache_lsh)
  saveRDS(section_scores, "cache/section-scores.rds")
} else {
  message("Loading LSH objects from cache.")
  load(file = cache_lsh)
}
```

How many pairs of matches do we have above `0.1` Jaccard score?

```{r}
section_scores %>% 
  filter(score > 0.1) %>% 
  nrow()
```