### 2015-11-04 022-visualizations-for-article.Rmd
Network graph with revised data

---
title: "Visualizations for article"
author: "Lincoln Mullen"
date: "November 3, 2015"
output: html_document
---

```{r message=FALSE}
library("readr")
library("igraph")
library("dplyr")
library("stringr")
library("parallel")
library("textreuse")
source("R/best-section-matches.R")
source("R/summarize-borrowings.R")
source("R/extract-date.R")
source("R/extract_code_names.R")
options("mc.cores" = 6L)
```

Read the data.

```{r}
cache_corpus <- "cache/corpus-sections-minash-n120-seed623.rda"
load(file = cache_corpus)
cache_lsh <- "cache/lsh-sections.rda"
load(file = cache_lsh)
borrowings <- section_scores %>% 
  rename(section_a = a, section_b = b, similarity = score) %>% 
  mutate(dissimilarity = 1 - similarity,
         code_a = extract_code_names(section_a),
         code_b = extract_code_names(section_b),
         year_a = extract_date(code_a),
         year_b = extract_date(code_b))
codes <- c(borrowings$code_a, borrowings$code_b) %>% unique()
```

Calculate section matches and summarize them for a single code.

```{r}
CA1851 <- best_section_matches("CA1851", borrowings, threshold = 0.2)
CA1851
summarize_borrowings(CA1851)
```

Apply that to all the codes.

```{r}
section_pct <- codes %>% 
  mclapply(best_section_matches, scores = borrowings, threshold = 0.2)  %>% 
  mclapply(summarize_borrowings) %>% 
  bind_rows()
section_pct
```

Do some filtering and munge into an edge list.

```{r}
# Note that original code is a misnomer here
edges <- section_pct %>% 
  mutate(original_date = extract_date(original_code),
         match_date = extract_date(match_code)) %>% 
  filter(percentage_sections >= 0.05,
         !is.na(match_code),
         original_code >= match_code) %>% 
  select(original_code, match_code, weight = percentage_sections)
edges
```

Turn that into a graph.

```{r}
g <- graph_from_data_frame(edges, directed = TRUE)
set.seed(4821)
l <- layout.fruchterman.reingold(g, weights = E(g)$weight)
par(mar = c(0,0,1,0))
plot(g, edge.width = E(g)$weight * 8, layout = l,
     edge.arrow.size = 0.65, vertex.size = 5)
title("Codes of Civil Procedure by percentage of sections")
```