### 2015-11-23 017-best-section-matches.Rmd
Updates to analysis

---
title: "Best matches for code sections"
author: "Lincoln Mullen"
date: "September 25, 2015"
---

```{r message=FALSE}
library("dplyr")
library("stringr")
library("readr")
library("tidyr")
library("ggplot2")
library("RColorBrewer")
source("R/extract-date.R")
source("R/extract_code_names.R")
```

The aim of this notebook is to create a function that takes a code's abbreviation (e.g., `NY1850`) and creates a data frame with each of the sections in that code with the best match to a non-anachronistic code.

Read in the data.

```{r}
load("cache/lsh-sections.rda")
scores <- section_scores %>% 
  rename(section_a = a, section_b = b) %>% 
  mutate(code_a = extract_code_names(section_a),
         code_b = extract_code_names(section_b),
         year_a = extract_date(code_a),
         year_b = extract_date(code_b)) %>% 
  rename(similarity = score) %>% 
  mutate(dissimilarity = 1 - similarity)
scores
```

```{r}
source("R/summarize-borrowings.R")
source("R/best-section-matches.R")
```

Apply this to three codes:

```{r}
CA1851 <- best_section_matches("CA1851", scores, threshold = 0.2)
CA1851
summarize_borrowings(CA1851)
IA1859 <- best_section_matches("IA1859", scores, threshold = 0.2)
IA1859
summarize_borrowings(IA1859)
UT1870 <- best_section_matches("UT1870", scores, threshold = 0.2)
UT1870
summarize_borrowings(UT1870)
```

Make a plot of code borrowings.

```{r}
plot_borrowings <- function(state_borrowings) {
  state_borrowings %>% 
    mutate(state = str_extract(match_code, "\\w\\w")) %>% 
    ggplot(aes(x = as.factor(all_sections), y = similarity, fill = state)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    scale_x_discrete(labels = NULL) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_brewer(palette = "Dark2")
}
```

Plot some borrowings:

```{r}
plot_borrowings(CA1851)
plot_borrowings(UT1870)
```

Heatmaps of borrowings:

```{r}
col_vector <- function(n, cols) {
  r <- trunc(n / cols)
  remainder <- n %% cols
  if (remainder > 0)
    c(rep(1:cols, r), 1:remainder)
  else
    c(rep(1:cols, r))
} 
row_vector <- function(n, cols) {
  out <- NULL
  for (i in seq_len((n / cols) + 1)) {
     out <- c(out, rep(i, cols))
  }
  out[1:n]
}
n_columns <- 40
```

Plot CA

```{r}
other <- c("MN1851", "IA1851", "KY1851", "MO1849")
best_section_matches("CA1851", scores, threshold = 0.3) %>% 
  mutate(match_code = ifelse(match_code %in% other,
                             "Other", match_code)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code, alpha)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in CA1851",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

Plot UT1859

```{r}
other <- c("NY1848", "WI1849", "CA1850", "NY1849", "IA1851", "NY1851", "WI1856",
           "WI1858", "NY1853", "CA1851", "CA1858", "IA1859", "IN1852", "MA1858",
           "MN1851", "TN1858", "KY1851", "MA1836", "MO1856", "NE1857", "OH1853",
           "TX1855", "KS1859", "NE1859")
best_section_matches("UT1859", scores, threshold = 0.3) %>% 
  mutate(match_code = ifelse(match_code %in% other,
                             "Other", match_code)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in UT1859",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

Plot WA1855

```{r}
white_list <- c("CA1850", "NY1850", "WI1849", "CA1851", "OR1854", "IN1852")
other_maker <- function(x, white_list) {
  ifelse(is.na(x), NA, ifelse(x %in% white_list, x, "Other"))
}
best_section_matches("WA1855", scores, threshold = 0.3) %>% 
  mutate(match_code = other_maker(match_code, white_list)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in WA1855",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

OR 1854

```{r}
white_list <- c("CA1850", "NY1853", "NY1851", "WI1849", "MN1851", "NY1850", "CA1851")
other_maker <- function(x, white_list) {
  ifelse(is.na(x), NA, ifelse(x %in% white_list, x, "Other"))
}
best_section_matches("OR1854", scores, threshold = 0.15) %>% 
  mutate(match_code = other_maker(match_code, white_list)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in OR1854",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

NC1868

```{r}
white_list <- c("DT1668", "NY1851", "NY1849", "NY1850", "WI1858", "WI1856")
other_maker <- function(x, white_list) {
  ifelse(is.na(x), NA, ifelse(x %in% white_list, x, "Other"))
}
best_section_matches("NC1868", scores, threshold = 0.3) %>% 
  mutate(match_code = other_maker(match_code, white_list)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in NC1868",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

IA1859

```{r}
white_list <- c("IA1851", "NE1857", "KY1854")
other_maker <- function(x, white_list) {
  ifelse(is.na(x), NA, ifelse(x %in% white_list, x, "Other"))
}
best_section_matches("IA1859", scores, threshold = 0.3) %>% 
  mutate(match_code = other_maker(match_code, white_list)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_code)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in IA1859",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

NM1897

```{r}
white_list <- c("MO", "NM", "IA", "NY", "MN", "WI")
other_maker <- function(x, white_list) {
  ifelse(is.na(x), NA, ifelse(x %in% white_list, x, "Other"))
}
best_section_matches("NM1897", scores, threshold = 0.15) %>% 
  mutate(match_state = str_sub(match_code, start = 1, end = 2)) %>% 
  mutate(match_state = other_maker(match_state, white_list)) %>% 
  mutate(., 
         column = col_vector(nrow(.), n_columns),
         row = row_vector(nrow(.), n_columns),
         match_code = as.factor(match_code)) %>% 
  ggplot(aes(x = column, y = -row, fill = match_state)) +
  geom_tile(color = "lightgray") + 
  theme_minimal(base_size = 16) +
  coord_equal() +
  labs(x = NULL, y = NULL, title = "Borrowed sections in NM1897",
       fill = "Section borrowed from") + 
  theme(legend.position = "bottom", 
        axis.ticks = element_blank(),
        axis.line = element_blank(), axis.text = element_blank(),
        panel.background = element_blank(), panel.grid = element_blank()) +
  scale_fill_brewer(type = "qual", na.value = "gray", palette = 6)
```

Summarize by state

```{r}
summarize_by_state <- function(section_list) {
  section_list %>% 
    mutate(match_state = str_sub(match_code, start = 1, end = 2)) %>% 
    group_by(match_state) %>% 
    summarize(original_code = unique(original_code),
              mean_similarity = mean(similarity),
              n = n()) %>% 
    mutate(percentage_sections = n / nrow(section_list)) %>% 
    arrange(desc(n)) %>% 
    select(original_code, match_state, mean_similarity, n, percentage_sections)
}
```