### 2015-11-23 024-comparison-NY.Rmd
Compare NY codes using NY1829

---
title: "Comparison of New York codes"
date: "November 23, 2015"
output: html_document
---

```{r}
library("textreuse")
library("dplyr")
library("readr")
library("stringr")
source("R/best-section-matches.R")
source("R/summarize-borrowings.R")
source("R/extract_code_names.R")
source("R/extract-date.R")
options("mc.cores" = 6L)
h <- 120
b <- 60
minhash <- minhash_generator(n = h, seed = 623)
ny_paths <- Sys.glob("legal-codes-split/NY*")
```

Create the corpus:

```{r}
ny_sections <- TextReuseCorpus(paths = ny_paths,
                               tokenizer = tokenize_ngrams, n = 5,
                               keep_tokens = FALSE,
                               minhash_func = minhash)
```

Do the LSH for New York:

```{r}
ny_buckets <- lsh(ny_sections, bands = b)
ny_scores <- ny_buckets %>% 
  lsh_candidates() %>% 
  lsh_compare(ny_sections, jaccard_similarity) %>% 
  rename(section_a = a, section_b = b, similarity = score) %>% 
  mutate(dissimilarity = 1 - similarity,
         code_a = extract_code_names(section_a),
         code_b = extract_code_names(section_b),
         year_a = extract_date(code_a),
         year_b = extract_date(code_b))
```

Write out the best section matches to disk.

```{r}
ny_codes <- c(ny_scores$code_a, ny_scores$code_b) %>% unique %>% sort

lapply(ny_codes, function(x) {
  best_section_matches(x, ny_scores, threshold = 0.15) %>% 
    write_csv(str_c("out/NY-comparisons-", x, ".csv"))
})
```

Summaries of matches.

```{r}
ny_codes %>% 
  lapply(function(x) {
    best_section_matches(x, ny_scores, threshold = 0.15) %>% 
      summarize_borrowings()
  })
```