### 2015-12-01 early-analysis/021-check-sectioning.Rmd

---
title: "Checking Sectioning of Codes"
author: "Lincoln Mullen"
date: "October 28, 2015"
output: html_document
---

Create a corpus of the sections.

```{r}
library("textreuse")
library("dplyr")
library("stringr")
library("readr")
options("mc.cores" = 6L)
sections <- TextReuseCorpus(dir = "legal-codes-split/", tokenizer = NULL)
```

Get the word counts of each section and turn that into a data frame.

```{r}
wc_raw <- wordcount(sections)
wc <- data_frame(section = names(wc_raw), words = wc_raw)
wc
```

Now detect the code name and calculate some summary statistics.

```{r}
wc_summary <- wc %>% 
  mutate(code = str_replace(section, "-\\d+", "")) %>% 
  group_by(code) %>% 
  summarize(n_sections = n(),
            total_words = sum(words),
            min_words = min(words),
            max_words = max(words),
            mean_words = mean(words),
            median_words = median(words)) %>% 
  arrange(code)

wc_summary %>% arrange(desc(n_sections))
wc_summary %>% arrange(n_sections)
```

And write that file to disk.

```{r}
write_csv(wc_summary, "out/section-wordcounts-summary.csv")
```