### 2015-12-02 031-count-sections.Rmd
Count sections in each code

---
title: "Number of actual sections"
author: "Lincoln Mullen"
date: "December 2, 2015"
---

Count the number of sections in each code.

```{r}
library("magrittr")
library("readr")
library("dplyr")
library("stringr")
library("knitr")
code_files <- Sys.glob("legal-codes/*")
codes <- lapply(code_files, read_lines)
names(codes) <- code_files %>% basename() %>% str_replace_all("\\.txt", "")
code_info <- read_csv("codes/code-generations.csv")
```

A section is a line with `§` or `SEC.` at the very beginning.

```{r}
pattern_clean <- regex("(^§|^SEC\\.)", ignore_case = FALSE)
counts <- codes %>% 
  lapply(str_detect, pattern_clean) %>% 
  sapply(sum, na.rm = TRUE)
section_counts <- data_frame(code = names(counts), sections = counts) %>% 
  left_join(code_info)
```

The number of sections in each generation.

```{r}
section_counts %>% 
  group_by(generation) %>% 
  summarise(sections = sum(sections)) %>% 
  kable()
```

The number of sections in each code.

```{r}
section_counts %>% 
  kable()
```