### 2015-12-03 index.Rmd
Add filename to index

---
title: "Field Code Analysis"
author: "Lincoln Mullen and Kellen Funk"
output: html_document
---

```{r echo=FALSE, results='asis'}
rmd <- Sys.glob("*.Rmd")
html <- sub(".Rmd", ".html", rmd)
lines <- lapply(rmd, readLines)
yaml <- lapply(lines, rmarkdown:::parse_yaml_front_matter)
cat("<ul>")
for (i in seq_along(rmd)) {
  cat(paste0("<li><a href='", html[i], "'>", yaml[[i]]$title, "</a><br/>",
             "<code>", html[i], "</code>", "</li>"))
}
cat("</ul>")
```