### 2015-12-03 index.Rmd
Infrastructure for web notebook

---
title: "Analysis of Text Reuse in the Field Code"
author: "Lincoln Mullen and Kellen Funk"
---

This is an analysis notebook for detecting text reuse in codes of civil procedure---the "Field codes"---in the nineteenth-century United States. It is a collaboration between [Kellen Funk](http://kellenfunk.org/) and [Lincoln Mullen](http://lincolnmullen.com).

See these links for more details:

- [Kellen's posts about the project](http://kellenfunk.org/field-code/)
- [An Introduction to the textreuse Package, with Suggested Applications](http://lincolnmullen.com/blog/an-introduction-to-the-textreuse-package/)
- [Detecting Text Reuse in Nineteenth-Century Legal Documents](http://lincolnmullen.com/blog/detecting-text-reuse-in-legal-documents/)

This notebook contains our work in progress. This is the most recent analysis in progress. There is more in the [repository](https://github.com/lmullen/civil-procedure-codes).

```{r echo=FALSE, results='asis'}
rmd <- Sys.glob("*.Rmd")
rmd <- rmd[rmd != "index.Rmd"]
html <- sub(".Rmd", ".html", rmd)
lines <- lapply(rmd, readLines)
yaml <- lapply(lines, rmarkdown:::parse_yaml_front_matter)
cat("<ul>")
for (i in seq_along(rmd)) {
  cat(paste0("<li><a href='", html[i], "'>", yaml[[i]]$title, "</a><br/>",
             "<code>", rmd[i], "</code>", "</li>"))
}
cat("</ul>")
```