### 2016-01-30 103-racial-exclusions.Rmd
Sections with racial exclusions

---
title: "Sections with Racial Exclusions"
author: "Lincoln Mullen"
output: html_document
---

I used `ag` to search the split sections for `negro OR indian OR china OR chinese OR mongol OR blood`. We are going to load the corpus and that list of files, the create a chronological list of those sections.

```{r}
library(readr)
library(textreuse)
library(stringr)
library(dplyr)
source("R/helper.R")
load("cache/corpus-lsh.rda")
```

Load the matching sections.

```{r}
exclusions <- read_lines("racial-exclusions/list-of-racial-sections.txt") %>% 
  str_sub(end = -5)
```

Get those sections in the right order.

```{r}
ordered_sections <- data_frame(section = exclusions,
                               year = extract_date(exclusions),
                               code = extract_code_names(exclusions)) %>% 
  arrange(year, code)
```

Print the sections to a file.

```{r}
get_printable_code <- function(doc_id) {
  require("stringr")
  content <- str_wrap(content(sections[[doc_id]]))
  str_c(doc_id, content, "-----------------------------------------------\n\n", 
        sep = "\n\n")
}

write_sections <- function(df) {
  require("stringr")
  docs <- df$section %>% sapply(get_printable_code) 
  text <- str_c(docs, collapse = "\n") %>% 
    str_split("\n") %>% 
    unlist()
  filename <- "racial-exclusions/exclusion-sections.txt"
  writeLines(text, filename)
  df
}

write_sections(ordered_sections)
```