### 2016-02-17 talks/dh-working-group-slides/dh-working-group-slides.Rmd
Updated slides for DH working group

---
title: "The Migration of the Field Code"
author: "Lincoln Mullen and Kellen Funk"
date: "http://lincolnmullen.com"
output: 
  revealjs::revealjs_presentation:
    theme: white
    css: talks.css
    transition: slide
    fig_width: 9
    fig_height: 6
    fig_caption: true
    pandoc_args: [
      "--slide-level=2",
      "--metadata=history:true",
      "--metadata=margin:0.05",
      "--metadata=center:false",
      "--variable=width:1200",
      "--variable=height:700"
    ]
    mathjax: null
    self_contained: true
    center: false
---

## The New York "Field Code"

<div class="left">
![David Dudley Field.](field.jpg)
</div>
<div class="right">
![Title page of the 1850 Field Code.](field-code.jpg)
</div>

## Code commissioners

![The 1868 North Carolina code commission: Victor Barringer, William Rodman, Albion Tourgee.](nc-commission.jpg)

## {data-background="field-code-states-map.jpeg"}

## "Paste-pot" procedure codes

![The 1861 Morris bill, which adapted California's code for Nevada.](morris-bill.jpg)

## {data-background="code-to-code2.png"}

## {data-background="state-to-state2.png"}

## {data-background="WA1855.png"}

## {data-background="CA1851.png"}

## {data-background="NV1861.png"}

## {data-background="NY1879.png"}