### 2016-02-19 article.Rmd
Move article template

---
title: 'The “Paste-Pot” Field Code: The Migration of Law in the Nineteenth-Century
  United States'
author: "Kellen Funk and Lincoln Mullen"
output: pdf_document
---

```{r setup, include=FALSE}
library(textreuse)
library(dplyr)
library(readr)
library(ggplot2)
library(ggraph)
knitr::opts_chunk$set(echo = FALSE)
# load("cache/corpus-lsh.rda")
# load("cache/network-graphs.rda")
# clusters <- read_csv("out/clusters/clusters.csv")
```