### 2016-03-15 106-innovative-sections.Rmd
Find innovative sections

---
title: "Innovations in codes"
author: "Lincoln Mullen"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The aim of this document is to figure out which sections were innovative (or at least, not borrowed) in a code.

```{r}
library(textreuse)
library(purrr)
library(dplyr)
library(readr)
load("cache/corpus-lsh.rda")
source("R/helper.R")
```

Function to get the non-matching sections.

```{r}
innovative_sections <- function(code_id, matches_df = best_matches) {
  require(dplyr)
  matches_df %>% 
    filter(borrower_code == code_id,
           is.na(match_section))
}
```

Test on a single code.

```{r}
innovative_sections("CA1851")
```

Functions to write the innovative sections to disk.

```{r}
write_sections <- function(sections_df, corpus = sections, filename = NULL) {
  require(stringr)
  if (nrow(sections_df) == 0) return(NA)
  corpus <- corpus[sections_df$borrower_section]

  texts <- lapply(corpus$documents, content) %>% 
    vapply(str_wrap, character(1))

  num_sections <- length(texts)
  code_name <- texts %>% names %>% extract_code_names() %>% unique %>% `[`(1)

  separator <- "------------------------------------------------------------\n"

  output_header <- str_c("Innovative sections (i.e., sections without best matches)\n",
                         "Code name: ", code_name, "\n",
                         "Number of sections: ", num_sections, "\n\n"
                         )

  output <- map2(names(texts), texts, function(name, text) {
    str_c("\n", separator, "\n\n",
          "Section ID: ", name, "\n\n",
          text, "\n\n")
  })

  output <- str_c(output, collapse = "")
  output <- str_c(output_header, output, collapse = "")

  if (is.null(filename)) filename <- str_c("innovations/", code_name, ".txt")

  message("Writing ", filename)
  writeLines(output, filename)

}
```

Write all the innovative sections to disk:

```{r}
unlink("innovations", recursive = TRUE)
dir.create("innovations", showWarnings = FALSE)
all_codes <- best_matches$borrower_code %>% unique %>% sort
all_codes %>% map(innovative_sections) %>% walk(write_sections)
```