### 2016-03-21 108-mapping.Rmd
Revert "Count number of sections for comparison to NY"
This reverts commit 04e65feb331d6ff53c08180debbcdb255e4a7cd2.

---
title: "Maps"
author: "Lincoln Mullen"
date: "March 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Make

```{r}
library(dplyr)
library(stringr)
library(textreuse)
library(readr)
load("cache/corpus-lsh.rda")
source("R/helper.R")
list1 <- read_lines("compare-to-NY-list-1.txt")
list2 <- read_lines("compare-to-NY-list-2.txt")
library(classInt)
```

```{r}
ny_codes <- c("NY1848", "NY1849", "NY1850", "NY1851", "NY1853")
map_matches <- all_matches %>% 
  filter(xor(borrower_code %in% ny_codes, match_code %in% ny_codes))

map_matches_ny_borrower <- map_matches %>% 
  filter(borrower_code %in% ny_codes) %>% 
  select(-borrower_state) %>% 
  mutate(match_section2 = borrower_section,
         borrower_section2 = match_section) %>% 
  select(-match_section, -borrower_section) %>% 
  rename(borrower_section = borrower_section2, match_section = match_section2) %>%
  select(borrower_section, match_section, score)

map_matches_not_ny <- map_matches %>% 
  filter(!(borrower_code %in% ny_codes)) %>% 
  select(borrower_section, match_section, score)

map_matches <- bind_rows(map_matches_ny_borrower, map_matches_not_ny) %>% 
  mutate(borrower_code = extract_code_names(borrower_section),
         borrower_year = extract_date(borrower_section),
         borrower_state = extract_state(borrower_section),
         match_code = extract_code_names(match_section),
         match_year = extract_date(match_section),
         match_state = extract_state(match_section))

map_matches <- map_matches %>% 
  filter(borrower_year >= match_year) %>% 
  group_by(borrower_section) %>% 
  arrange(desc(score)) %>% 
  slice(1) %>% 
  ungroup() %>%  
  filter(score >= 0.2)

map_matches <- data_frame(borrower_section = names(sections)) %>% 
  left_join(map_matches, by = "borrower_section") %>% 
  mutate(score = ifelse(is.na(score), 0, score))

map_summary <- map_matches %>% 
  filter(borrower_code %in% list2) %>% # Just the codes Kellen picked
  group_by(borrower_state) %>% 
  summarize(score = mean(score)) %>% 
  arrange(desc(score))

intervals <- classIntervals(map_summary$score, n = 5, style = "jenks")
map_summary$score <- cut(map_summary$score, breaks = intervals$brks)
```

```{r}
library(ggplot2)
library(broom)
library(USAboundaries)
library(rgeos)
library(maptools)
library(scales)
library(RColorBrewer)
us <- us_states()
us_fortified <- tidy(us, region = "stusps") %>% 
  left_join(map_summary, by = c("id" = "borrower_state")) %>% 
  filter(id != "AK",
         id != "HI",
         id != "PR")

ggplot() +
  geom_polygon(data = us_fortified,
               aes(x = long, y = lat, group = group, fill = score),
               color = "black", size = 0.25) +
  scale_fill_manual("Avg. dissimilarity score", values = brewer.pal(5, "YlGn")) +
  coord_map("albers", lat0=30, lat1=40) +
  theme_classic() +
  labs(title = "Strength of borrowings from Field Code (List 2)",
       x = NULL, y = NULL) +
  theme(axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "bottom")
```

```{r}
ggplot(map_summary, aes(x = borrower_state, y = score)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  coord_flip()
```