### 2016-05-27 104-network-graphs-in-ggraph.Rmd
Updates to network graph

---
title: "Network graphs in ggraph"
author: "Lincoln Mullen"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(igraph)
library(ggplot2)
library(ggraph)
library(stringr)
library(purrr)
load("cache/corpus-lsh.rda")
load("cache/network-graphs.rda")
source("R/helper.R")
set.seed(82892)
```

The network graphs aren't very pretty so far, so we are going to see if we can do better in [ggraph](https://github.com/thomasp85/ggraph).

```{r fig.width=12, fig.height=12}
ggraph(codes_g, "igraph", algorithm = "fr") +
  geom_edge_fan(aes(edge_width = sections_borrowed, 
                    alpha = sections_borrowed),
                arrow = arrow(type = "closed", ends = "first",
                              length = unit(0.20, "inches"),
                              angle = 15)) +
  geom_node_point(aes(colour = as.factor(distance)), size = 6) +
  scale_edge_width("Sections borrowed", range = c(1, 2), guide = "none") + 
  scale_edge_alpha(range = c(0.3, 0.6), guide = "none") +
  scale_color_discrete("Distance from a NY code") +
  ggforce::theme_no_axes() +
  geom_node_text(aes(label = name)) +
  labs(title = "Code to code borrowings") +
  theme(legend.position = "bottom",
        panel.border = element_blank())
```

Now make a plot of state to state borrowings.

```{r}
ggraph(states_g, "igraph", algorithm = "nicely") +
  geom_edge_fan(aes(edge_width = sections_borrowed, 
                    alpha = sections_borrowed),
                arrow = arrow(type = "closed", ends = "first",
                              length = unit(0.20, "inches"),
                              angle = 15)) +
  geom_node_point(size = 6, alpha = 0.75, aes(color = region)) +
  geom_node_text(aes(label = name)) +
  scale_edge_width("Sections borrowed", range = c(1, 2), guide = "none") + 
  scale_edge_alpha(range = c(0.3, 0.4), guide = "none") +
  scale_color_manual(values = c("South" = "steelblue1",
                                "West" = "coral2",
                                "Atlantic" = "darkgoldenrod2",
                                "New England" = "mediumorchid2",
                                "Midwest" = "palegreen2",
                                "New York" = "peru")) + 
  ggforce::theme_no_axes() +
  labs(title = "State to state borrowings") +
  theme(legend.position = "bottom",
        panel.border = element_blank())
```

Get section-to-section network graphs.

```{r}
library(igraph)
sections_edge_list <- best_matches %>% 
  select(borrower_section, match_section, score) %>% 
  filter(!is.na(match_section))
sections_g <- graph_from_data_frame(sections_edge_list, directed = TRUE)
sections_g <- decompose(sections_g, min.vertices = 5)
sections_v_count <- sapply(sections_g, vcount) 
hist(sections_v_count,main = "Number of vertices in section to section graphs")

section_of_interest <- "NY1850-006060"
has_section <- function(x, section) section %in% x
sections_g %>% 
  lapply(vertex_attr, "name") %>% 
  sapply(has_section, section_of_interest) %>% 
  which()

# 484 for witness testimony
# 1 for abolishing chancery
```

```{r}
plot_graph <- function(g, title = "A cluster of borrowings", subtitle = NULL) {
plot <- ggraph(g, "igraph", algorithm = "nicely") +
  geom_edge_fan(aes(edge_width = score, 
                    alpha = score),
                arrow = arrow(type = "closed", ends = "first",
                              length = unit(0.10, "inches"),
                              angle = 15)) +
  geom_node_point(size = 6, alpha = 0.5, color = "lightblue") +
  geom_node_text(aes(label = name)) +
  scale_edge_width("Similarity score", range = c(1, 2), guide = "none") + 
  scale_edge_alpha(range = c(0.3, 0.4), guide = "none") +
  ggforce::theme_no_axes() +
  labs(title = title,
       subtitle = subtitle) +
  theme(legend.position = "bottom",
        panel.border = element_blank())
}

set.seed(100)
abolish_chancery <- sections_g[[1]]
V(abolish_chancery)$name <- str_replace_all(V(abolish_chancery)$name, "-\\d+", "")
civil_action <- plot_graph(abolish_chancery, title = 'The Creation of "A Civil Action"')
ggsave("out/civil-action-sections-network.png", civil_action, width = 8, height = 6)
```