### 2016-06-26 article/article.Rmd
Cite all the text reuse and repository things

---
author:
- Kellen Funk
- Lincoln Mullen
fontsize: 12pt
output:
  pdf_document:
    fig_caption: True
    md_extensions: '-tex_math_single_backslash'
    number_sections: False
title: The Spine of American Legal Practice
---

At the opening of the first Nevada legislature in 1861, Territorial Governor James W. Nye, a former New York lawyer, instructed the assembly that they would have to forsake the Mormon statutes of Utah Territory from which Nevada had been carved. Those laws were ill adapted to "the mining interests" of the new territory, but "happily for us, a neighboring State whose interests are similar to ours, has established a code of laws" sufficiently attractive to "capital from abroad." That neighbor, of course, was California, and Nye urged that California's code of civil practice be enacted in Nevada, as far as it could "be made applicable."[^1] Territorial Senator William Morris Stewart, a famed silver-mining lawyer who would become the leading U.S. senator of Reconstruction, followed the instructions perhaps too well. Stewart literally cut and pasted the latest *Wood's Digest* of the California Practice Act into the session bill, crossing out *state* and *California* and substituting *territory* and *Nevada* where necessary. Stewart copied not just California's procedure code but also its method of codification, for California had in turn borrowed its code by modifying New York's.

![Detail from Council Bill 21, First Territorial Legislative Session (1861), Nevada State Library, Archives and Public Records.](morris-bill.jpg)

Nye wrote back to the assembly in disgust. The bill---of 715 sections---had reached him late the night before the legislative session was to close. Even in the few hours he had to read it, Nye counted "many errors in the enrolling of it, numbering probably more than three hundred." Some errors were severe. Nevada's Organic Act specified the jurisdiction of the district courts and justices of the peace, but the code overwrote these limitations by copying California's arrangements. Error-riddled and unconstitutional as the bill was, Nye believed a civil practice code was a "universal necessity and public need," concluding that "without some legislation in providing a code of procedure in civil cases, it may be a matter of grave question whether your courts would be able to fulfill the purpose of their creation."[^2] Nye signed the code into law.

Strikingly, nothing like this "universal necessity" for a code of practice existed when Nye began his legal career in New York in the 1840s; rather, it was one of the central developments of America law during the nineteenth-century. By the end of the nineteenth-century, thirty-one American states and territories had adapted the text of a New York code of civil procedure first promulgated in 1848. The code colloquially became known as the Field Code, named for its chief draftsman David Dudley Field, a famed Manhattan trial lawyer.[^3] When Field's code appeared in the Colorado assembly in 1876, a Denver newspaper wryly commented, "The scissors and paste-pot we had heretofore confidently believed were implements peculiar to the newspaper sanctum." To the editor, the cut-and-paste code was not just a humorous curiosity of legislative history. Rather, the extensive copying of Field's Code threatened the foundations of American popular sovereignty. "The bill is a long one; the assembly has not the time to devote to it and to give it the scrutiny that a measure of such importance demands," the editor explained. Blind approval "would be an injustice to themselves and a greater wrong and injustice to the people who have a right to demand that their public servants legislate for the public good." The code "has been 'assimilated,' as we are informed, 'to the character and requirements of our people,' whatever that may mean," but the editor feared the legislation was the product of "men who have the welfare of the 'dear people' at their tongue's end always, but never in their hearts."[^4]

The migration of the Field Code was a central event in Anglo-American legal history, but no modern historian has traced the extensive borrowings of the Field text nor recognized the political furor that greeted the code outside New York.[^5] Every aspect of a civil justice system, from the rules granting access to courts, to lawyers, to remedies (whether damages, injunctions, or seizure of property) was covered by the code, making its New York-specific rules politically contentious both inside and outside the Empire State. As the Field Code migrated around the country, commentators in each jurisdiction raised the same complaint: how could legislation borrowed from another state represent the popular will and bests interests of *this* state?

Understanding the history of the Field Code requires not only attention to its political context but also a detailed examination of the substance of what was borrowed and what was revised in each jurisdiction. Exploring these borrowings is a daunting task, however. Procedure codes were long, technical documents, and although each jurisdiction copied large swaths of text, each also modified the text along the way, sometimes with a simple *Nevada* for *California*, sometimes with more foundational changes to the remedial system. Although Stewart's cut-and-paste code found its way into the archives, most draft legislation did not, and few codifiers explained in detail how they produced their texts. Traditional reading or textual criticism of some 98,000 distinct sections of law comprising some 7.7 million words is simply not a feasible research task for a historian who wishes to track these borrowings.

Yet by turning to the digital analysis of texts, we have resolved this difficulty and tracked the how the sections of the various codes of civil procedures were borrowed from one another. Within the corpus of legislation, algorithmic analysis of texts can essentially reverse engineer and visualize what the archive revealed in Figure 1: which texts were borrowed, which modified, and how extensively. Our method works especially well for legal texts, for reasons we will explain. However, scholars can extend this method to other kinds of curated collection of texts, from church hymns to medical treatises.

The first contribution of this article is the demonstration of our methods as applied to the corpus of nineteenth-century state procedure codes. The second contribution is the integration of what we learned from the text analysis with more conventional approaches of political and cultural history to explain the importance of the migration of the Field Code. On the national level the extent of legislative borrowing followed a pattern American historians have described as a "Greater Reconstruction" in which the former Confederate South and the Far West showed a remarkable kinship. Scholars have typically described Greater Reconstruction as a federal development, featuring the creation of national citizenship, a national economy, and a larger state apparatus centered in Washington, D.C. This article shows that Greater Reconstruction had its state-level dimensions as well. The uniform practice of law and adjudication of civil remedies was not structured by Washington mandates, however, but by the anxiety that New York financial capital would follow only New York civil remedies. At the more local level, our digital computations can trace modifications within code traditions, for instance, the ways western and midwestern codifiers altered New York's law to accommodate hardening conceptions of racial competencies in the civil courts. To see these developments, one needs a tool that can canvass all the states at once across broad swaths of civil law.

# 1. The origin and migration of the Field Code

Until 1848, civil remedies and trial practice in New York were largely governed by common law traditions loosely categorized as "practice and pleadings."[^6] To understand how to file a civil claim or to enforce a judgment, a lawyer had to consult ad hoc statutes from colonial times to the present as well as precedents reported from cases litigated at common law and chancery. By the 1840s, enterprising practitioners had collated these materials into a half dozen marketable treatises, but these remained works of private opinion---no court was bound to agree with the treatise writers as to the weight, relevance, or proper interpretation of a legal statement.[^7] The common law was accordingly known as "unwritten law" despite the proliferation of published texts. The point was that common law was not precisely determined until a particular case drawn from the messy facts of life demanded resolution.[^8] Statutes, on the other hand, were "written law," prescribing or reforming the rules even before a case put the precise question in issue. And within the realm of written law, codes were the ultimate statutes.[^9]

Codification is, as Lawrence Friedman has written, "one of the set pieces of American legal history." Law reformers advocated for a codification of the common law from the earliest days of the Republic up into the Gilded Age, from Massachusetts down to South Carolina. Efforts ranged from mere compilations of existing statutes in each state to a full European-style codification meant to be an entirely comprehensive and systematic statement of the law.[^10] At its most basic level, codification proposed that legislative policy ought to be the sole source of law. Law was to be made by democratically responsible legislators---in terse, unambiguous statements---not discovered through application and analogy in particular cases by the judges. Debates over codification thus ranged from the metaphysics of law to political theories of separation of powers and institutional competency.[^11]

In New York, these codification debates came to a head at the 1846 constitutional convention, where "the conquerors took all," as the ambivalent Jacksonian James Fenimore Cooper complained.[^12] Law reformers abolished the court of chancery, made judges stand for popular election, and required the legislature to appoint a commission to reform the "practice and pleadings" of the civil courts.[^13] New Yorkers had two models of legislative commissions on which they could draw for their own law reforms. The French government under Napoleon had appointed five-member commissions to codify the law of France. When the New York law reformer William Sampson called for codification in a widely noted address to the New York historical society, one of the members of the French commission was living in exile in upstate New York at the time and wrote to Sampson to advise him on the mechanics of codification: "Let four or five good heads be united in a commission, to frame in silence the project of a code. It is not so difficult a task. It is only to consult together, and to select. Do so with your best authors as we did with ours, ... which we simply converted into articles of our code." Tellingly, the French commissioner took it as granted that the commission's code would automatically be promulgated as law.[^14]

The other model came from England. Royal commissions had been employed since before the Revolution of 1688 to advise on a variety of matters. After the Revolution, royal commissions became disfavored while Parliament undertook many of the same functions through "committees of enquiry" to which only members of Parliament were appointed. By the nineteenth century, royal commissions again became common. Royal commissioners were usually not legislators. On the one hand, that meant commissioners were not bound to the rhythms of Parliament's schedule---they could continue their work even as Parliament adjourned. On the other hand, the recommendations of royal commissioners were a further remove from actual legislation. Although by the mid-nineteenth century royal commissions sometimes offered model statutes, the legislation they advised still had to pass through the normal politicking and drafting processes of Parliament.[^15]

After David Dudley Field and two other lawyers were appointed to the procedural commission, their reports made clear that they favored the French model but understood political realities would hold them to the English model. From 1847 to 1850, the commissioners made five reports to the legislature knowing they had no power to keep legislators from amending their code or even defeating it altogether. Each time they reminded legislators that "public opinion had issued its mandate in the most imposing form" of a constitutional decree. The constitution, together with the legislative act appointing the commission, "gave the commissioners instructions so precise, as to leave them no discretion, if they had desired it, \[and\] promised them therefore in advance, so long as they obeyed those instructions, the concurrence and co-operation of all departments of the government."[^16] Although the theory of codification made it a democratic enterprise, in practice Field---an ardent Jacksonian---insisted the democratic legislature ought to defer to the expertise of the commissioners.

The Field commission sought to blunt criticism by insisting that political concerns about lawmaking did not apply to mere procedure. "The system of procedure by which law is administered, differs from the law itself in this," the commissioners explained: "the latter is a body of elementary rules founded in the immutable principles of justice, drawing their origin from the obligations which divine wisdom has imposed ...; while the former consists, in its very nature, but of a body of prescribed rules, having no source but the will of those by whom they are laid down." Substantive law was universal, natural, grounded in divine justice, and therefore entitled to respect and protection from change. But God cared nothing of the "the mere machinery by which law is to be administered." Thus, the commissioners argued, procedure was trivial enough for legislative experimentation but complicated enough that only master practitioners like themselves could run the experiment.[^17]

Yet the scope of the procedure Code included far more than the "mere machinery" of a lawsuit.[^18] The final draft of the code, printed in 1850, spanned nearly 800 pages of 1,885 regulations. The first third of the code covered constitutional topics, specifying the jurisdiction of all state courts and the duties of all state officers (and liabilities for violating those duties). The code deregulated attorney compensation, introducing novel structures of retainers and contingency fees.[^19] It created summary procedures meant to accelerate debt collections while simultaneously carving out "homestead" exemptions from the sheriff's reach.[^20] The code concluded by defining who could be an attorney, a juror, and a witness, drawing racial and gendered distinctions over who could speak in court.[^21]

Most important, the code defined all the remedies that a civil court could order---from money damages, to partition of property, to injunctive decrees and contempts---and made those remedies available in every lawsuit. In many cases, a legal right was indistinguishable from the remedy that secured that right: the right to possess a particular piece of property and the remedy that seized and delivered that property were, in effect, the same thing. Remedies were thus intimately connected to substantive law. For that reason, neither the French *code de procédure civile* (1806) nor Blackstone's or Bentham's writings conceived of remedies as purely procedural.[^22] By codifying remedies, Field invited continual expansion of the category of "procedure." As other states adopted the procedure code, they sometimes included other fields of law that seemed obviously "substantive" yet had such specific procedures or remedies that they were placed in a "code of procedure," including the law of wills, corporations, and mortgages.[^23] After all, argued procedural codifiers in Iowa, what did the famed Married Women's Property Acts offer besides procedural reform? These acts gave women *standing* to litigate in their own name and seek *remedies* in claims of property and contract, and they abolished mandatory rules of *joinder* (of husbands). Thus, one of the most significant changes to the law of property and domestic relations in the century went into the state's Code of Civil Procedure.[^24]

Despite Field's arguments that a mere procedure code was democratically unproblematic, his efforts were not entirely successful in New York. The commission submitted a draft of its main reforms in 1848, emphasizing that this first code was "but a report in part." New York's legislators enacted the partial code with little amendment, some legislators repeating Field's view that the constitution obligated them to accept the code.[^25] But when the commissioners submitted their final draft in 1849, the Assembly judiciary committee balked, directly disputing the commissioners' claims that procedure was merely the machinery of the law. They wrote that "provisions for rights and for the mode of pursuing remedies, insensibly run into each other," and observed that distinctions between substance and procedure were difficult to draw in practice. In fact, "the parts of a system of law are more dependent on each other, and run into and affect each other, infinitely more than any machine of human contrivance." They therefore suspected the commissioners' forthcoming code of criminal procedure would include all of the criminal law as well, "as they seem to understand practice and pleadings to include all the law upon a given subject." That being the case, the committee wondered whether they should "place in \[the commissioners\] a blind and implicit confidence that shall commit to their discretion the peace and property, the personal liberty and the lives of those who sent *us* here to make laws for them?"[^26]

![This map shows which states adopted codes of civil procedure based on the New York Field Code. The date shown is the date of the first enactment of a procedure code; most states subsequently revised their codes. Note that many southern states and western states came to adopt the Field Code during the period of Greater Reconstruction. By the end of the nineteenth-century, thirty-one jurisdictions (those displayed on the map, plus Alaska) had adopted a version of the Field Code.](field-code-states-map.jpeg)

The code would encounter similar difficulties in each jurisdiction that adopted it. Even the shortest version of the Field Code was significantly longer than any other state statute before the Progressive legislation of the twentieth century. And unlike statutory compilations that sometimes took the name of a "code" but made no changes to existing law, the Field Code opened by abolishing the hallmarks of prior practice and instituting "hereafter" a new form of action with substantial revisions to basic matters of standing, remedy, and enforcement of civil judgments.[^27] In the states where it was imported, there was be no getting around the fact that the code introduced much new law, yet legislators were unable to read, critique, and amend the code within the brief period of a legislative session. "It is folly to undertake to pass a code in a sixty day session," wrote the *Montana Post*, "and the best way would be for the Assembly to select one from a State or Territory which would come near meeting our wants, and slide it through with the fewest changes possible."[^28] Sliding the code through eased the problem of time but exacerbated the problem of local sovereignty. "To be governed by a foreign law, especially when that law is not preknown to the people whose conduct is to be regulated thereby ... is something repugnant to the idea of Democratic Republican government," complained the *The Miner's Express* in Iowa.[^29]

How, then, did states and territories achieve a politically acceptable balance between efficiency and sovereignty, borrowing law for sake of time but endowing it with popular legitimacy in each locale? Quite apart from the technicalities of legal practice, the American federation of civil government into (depending on the year) more than thirty or forty separate state jurisdictions makes it hard to describe a phenomenon that was truly national despite its state-centered enactments. It requires a sense of how much law was borrowed in each location and to what degree innovations were introduced. But precisely because these questions concern codes---texts that comprehensively and systematically cover a given subject---they are ideal sources for the techniques of digital history.

# 2. Detecting borrowings among procedure codes

To discover how the Field Code migrated to other jurisdictions, we compiled a corpus of relevant (or possibly relevant) laws, including separately bound codes of civil procedure as well as codes or regulations appearing as session laws and statutory compilations. The corpus comprises 135 statutes from the nineteenth century, which amounts to 7.7 million words organized into 98,000 regulations. It includes the initial enactment of every code of civil procedure, as well as procedure statutes and re-enacted codes from states reputed to have been legally influential, but it does not include every known statute of American procedural law. While a comprehensive project may be illuminating in its own ways, our specific question of how New York legislation influenced other states permits a more curated corpus.[^30]

Curating a corpus to answer a specific question is an important difference from the way that digital history often describes itself. Speaking in broad terms, digital historical work---like all historical work---can begin either with a set of sources or with a question. Digital history is often conceived of as beginning with very large datasets such as the Google Books or Hathi Trust corpora. These large corpora are sometimes called "big data"---though, it must be emphasized, almost never by digital historians who actually work with them---and are used as the basis of a "distant reading." We think that working with these large corpora is worthwhile because it is salutary for historians to have their research questions shaped by the broadest possible contexts. Nevertheless, it is not apparent that digital historians can readily move from these omnibus corpora to answering the specific research questions that animate various historical fields. In this article we demonstrate an alternative approach, which we might unimaginatively label "medium data." Although the amount of legislation that governed American civil practice is impressive---every state amended and re-enacted procedure statutes nearly every decade---a corpus of procedural legislation requires some computational sophistication but is far from needing the computational techniques derived for truly big data. We have gathered a large but narrowly constrained corpus centered on solving a well-defined research question. This corpus is large enough that digital historical methods provide results that a scholar could not obtain through traditional methods, but sufficiently circumscribed so as to directly address a discipline- and field-specific question.[^31]

Because most codes were public statutes, they were widely printed and distributed and therefore found their way into libraries digitized by Google Books. We drew primarily from the Google Books, filling in gaps from other databases as necessary.[^32] We used optical character recognition software (OCR) to create plain-text versions of the codes, which we edited lightly, correcting section markers by hand as necessary and writing a script to fix the most obvious OCR errors.[^33]

The most important step we took in processing the files was to split each section of the code into its own text file. Codes varied in how they were organized, but they universally divided specific regulations into *sections* (or, on occasion, *articles*). Not only does the discursive form of these texts provide a handy organizational scheme for digital methods, but historically sections were also the way legislators borrowed their texts. Codifiers took their sources apart by sections, rearranging here, editing, drafting, and then re-combining there. Despite the fact that states differed widely on what topics they included in "civil procedure," sectioning the codes allowed us to assess similarity even among codes of quite different lengths and coverage. For instance, we know that California's 1851 code was derived from New York's 1850 code. (Stephen Johnson Field, David Dudley Field's brother, was the lawyer who imported New York's code into California.[^34]) But the New York code is over 150,000 words long, whereas California's code was just over 50,000 words long. Those disparate lengths mean that comparing all of the California code to all of the New York code is a less meaningful comparison that than comparing each section in the California code to each section in the New York code, where matching sections will have a similar length.

Having divided the texts according to a historically justified pattern, our next step was to compare each section to every other section and measure the similarity between them. To continue the New York-to-California example, consider the following pairs of sections. The first pair is from the final draft of the New York Field Code. These sections completely abolished prior practice and began to rebuild the procedure system from the ground up:

> § 554. The distinction between actions at law and suits in equity, and the forms of all such actions and suits, heretofore existing, are abolished; and, there shall be in this state, hereafter, but one form of civil action, for the enforcement or protection of private rights and the redress or prevention of private wrongs, which shall be denominated a civil action. § 555. In such action, the party complaining is known as the plaintiff, and the adverse party as the defendant.

In the theory of Euro-American lawyers, California had no prior practice to abolish, so the code began more simply:[^35]

> § 1. There shall be in this State but one form of Civil Action, for the enforcement or protection of private rights, and the redress or prevention of private wrongs. §2. In such action the party complaining shall be known as the plaintiff, and the adverse party as the defendant.

The pairs are obviously related to one another, both in terms of their legal force and in terms of the actual words used.

A common method for measuring the similarity of two documents involves dividing texts up into tokens of consecutive words (called n-grams) and calculating a Jaccard similarity score, defined as ratio between the number of tokens that the two document have in common to the total number of tokens that appear in both documents. We used five-word tokens and "shingled" them, meaning that for the New York sections above, the first token was "the distinction between actions at," the second token was "distinction between actions at law," and so on. These tokens each contain more meaning than a single word, yet because they are shingled they are robust to changes in the text or noisy OCR. A Jaccard similarity score will always be in a range between 0 (complete dissimilarity) and 1 (complete similarity).[^36]

Comparing the section pairs above produces the following matrix of scores.

                NY1850-554   NY1850-555   CA1851-001   CA1851-002
  ------------ ------------ ------------ ------------ ------------
  NY1850-554                     0           0.14          0
  NY1850-555                                  0           0.41
  CA1851-001                                               0
  CA1851-002

As we would expect the first respective sections (New York §554 and California §1) have a score of 0.14, which indicates that they are similar but have significant differences, while the second sections (New York §555 to California §2) have a much higher similarity score since only a few words were changed. Just as important, when we compare the first section in New York to the second section in California, we get a score of 0; the two sections are nothing like each other.

The aim, then, was to create a triangular matrix like the one above, but with approximately 98,000 rows and 98,000 columns, containing the similarity scores for each possible pair of sections. While this is easy enough to conceptualize, such a matrix is actually quite large, containing about 4.8 billion comparisons. This would take an unreasonable amount of computation time, and most of these comparisons would be unnecessary, since each section has no relationship to most other sections. We therefore implemented the minhash/locality sensitive hashing (LSH) algorithm to detect pairs of possible matches quickly.This algorithm works by extracting a set number of random tokens from each document to create a document signature; this subset makes it possible to approximate the Jaccard similarity of two documents. These signatures can then be used to identify each of the billions of possible pairs of sections as likely to be matches or not. By setting parameters to the algorithm, one can control how sensitive the algorithm is in drawing the line between genuine and spurious matches. We then measured the actual Jaccard score for only those candidate pairs.[^37]

The result was a matrix of similarity scores, with over 45,000 genuine matches. For each section in the corpus, after making some adjustments to remove anachronisms and spurious matches, we were able to identify the section from which it was most likely borrowed.[^38] In other words, we had traced the work of the commissioners' scissors and paste-pots through the course of their codes.[^39]

# Part 3

Having developed a method for finding legislative borrowings, we next sought to visualize the broader patterns revealed by the those borrowings. We did this first by aggregating the section to section borrowings into code to code borrowings. By counting how many sections each code borrowed from each other code, we were able to show the connections from one code to another.

The resulting network graph (**Figure X: Network Graph**) tells the genealogy of civil procedure in the United States. The New York codes, especially the finished draft of 1850, are central to the entire network.[^40] Yet it is important to note that New York spurred different traditions within the procedural network. Variations in the Field drafts meant that different states could borrow different versions of the Field Code. The 1850 New York code was the primary progenitor of several families of codes in California, Kentucky, Iowa, and Ohio, each of which in turn became major contributors to the law of neighboring states. The 1851 New York code became the progenitor of codes for Wisconsin, Florida, North Carolina, and South Carolina. Mississippi and Missouri borrowed from the New York 1849 code. These variations point out how the Field Code was always a moving target. While Field considered the 1850 New York Code to be the definitive, ideal version of the code, it was really all of the New York codes from 1848 to 1853 that became models for other jurisdictions, likely due to whatever material a given commission had on hand. The Field Code was not a single volume on the shelf, but a series of volumes any one of which might be more accessible in different regions and different eras.

Even later New York codes can be considered a separate family. In 1876 a New York commission produced a new code attempting to consolidate all the statutory amendments subsequent to the 1851 Field Code. David Dudley Field was upset by the changes introduced in this revised code, called the Throop Code. A count by a "friend" of Field's found that only three sentences of the Field Code had carried over word-for-word into the Throop Code. Nevertheless, many of the changes were purely formal, and Field speculated that these changes had been made only "so that \[the code\] should not appear to be the same thing as before." With respect to Mr. Field or his "friend," we were able to find that the connection was between the Throop Code and the Field Code was somewhat stronger than he thought, and the Throop Code was indeed a descendant, if much revised, from the Field Code.

Finally, our corpus included a number of statutes which stood outside the Field Code tradition, such as Virginia and West Virginia regulations, statutes from Massachusetts and Maine, and southern codes from Georgia to Louisiana. These statutes show that the dominance of the Field Code was not total, and a number of older jurisdictions remained outside of its ambit.[^41] But virtually every jurisdiction established or reconstructed after 1850 became a part of the Field Code network, and no other tradition achieved anywhere near the same coherence across state lines.

We have so far discussed code to code borrowings which give us a macroscopic view of the corpus as a whole. But because the fundamental unit of comparison is section to section, it is possible to get a more detailed look at the kinds of borrowings that codes made of one another, by looking at each section in a code and seeing which other code it was borrowed from. These visualizations show how each code was put together. Codes at different removes from New York tended to have different patterns of borrowing. To illustrate this, we will follow one branch of the Field Code network, beginning with the family started by California's 1851 code.

<!-- ![](media/image04.png){width="6.5in" height="3.7916666666666665in"} -->
California's 1851 code borrowed heavily from two sources. The bulk of the code was taken directly from New York's 1849 or 1850 codes (which are difficult to distinguish exactly from one another.[^42] The remainder of the code was borrowed from California's 1850 code. The 1850 version was in turn borrowed from the New York codes **TODO: might as well show this, esp. since CA1850 is so impressive**. As we would expect, the bulk of the borrowings that come from California instead of directly from New York are at the beginning of the code, where the 1850 code had made amendments specific to the borrowing from California. (Many of the non-matching sections are tables of contents.) The main point is that that the majority of the code was lifted more or less directly from New York, with minor amendments specific to California. California made few to no innovations to the Field Code beyond a rearrangement of its provisions.

<!-- ![](media/image05.png){width="6.5in" height="4.291666666666667in"} -->
The pattern of borrowings in the Washington 1855 code was a rather different case. The Washington code was definitely in the lineage of the 1851 California code, since it borrowed sections from both California directly and from Oregon, which was derived from California. While there are a number of "other" sections, the pattern of borrowing is clear: Indiana's 1852 code and Oregon's 1854 code provide the majority of the borrowings. Those two contiguous bands of borrowings correspond to regulations on judgement from Oregon and enforcement provisions from Indiana. This pattern likely came about because one of the Washington code commissioners, Edward Lander, was an Indiana appellate judge from 1850 to 1853, while another commissioner, William Strong, was a justice of the Oregon Supreme Court in the same years. While working on the Washington code, they must have each used the law (and the law books) that they knew best. As a second generation variation on a regional code, the Washington code drew from different sources.

Finally, we can consider the case of Washington's revised 1873 code as an example of one of the last "leaves" on the family tree of the Field Code. The bulk of this code was lifted from the Washington 1855 code with only small amendments. The main exception was a probate code drawn from California's 1872 code.[^43] Like many of the last generation codes, the text of the procedure code had stabilized in a local manifestation of a regional tradition. The code was still genuinely a Field Code, with a great deal of similarity to the original New York Field Codes, but its specific form depended on the many edits and rearrangements that state code commissioners had made to the text.

<!-- ![](media/image03.png){width="6.5in" height="4.666666666666667in"} -->
So far we have retained the context of the surrounding sections within a particular code. But since our fundamental unit of comparison is section to section, we can use a technique called clustering to deform the texts (as a literary scholar might say) in order to see the textual history of how particular regulations embedded in individual sections changed over time.[^44]

We used a clustering algorithm to group related sections together. Clustering is the process of taking a set of objects (in this case, sections of the field codes) which have distances between one another (in this case, the Jaccard similarity is easily converted into a distance measure), and grouping items which are are most alike together. There are innumerable clustering algorithms, but we needed one that could work with a sparse matrix and one whose assumptions matched the characteristics of our problem. We decided to use affinity propagation clustering.[^45] That algorithm assumes that each cluster has an "exemplar" item which the other items are similar to. That assumption fits nicely with borrowings from the Field Code, where a single section (likely from the 1850 New York code) had many borrowings, but where there could also be innovative sections in other stages.[^46]

The result was a set of approximately 2,900 clusters which contained at least five sections, though this probably overstates the number of innovative, *ur*-sections in the corpus. Each cluster contained a list of the section that belonged to it. The biggest cluster, for instance, which concerned the use of affidavits in pleading, contained 103 sections. With those clusters, we were able to organize the documents chronologically.[^47] This method is particularly useful when tracking the variations in in the law from jurisdiction to jurisdiction. It provides historians with a way of noticing small changes in the wording and substance of the law. Most discussions of algorithmic reading have focused on "distant reading,"[^48] or have balanced the claims of distant reading by using it as a means to enable close reading. This method of clustering, however, is a kind of algorithmic close reading. By deforming the texts---taking them out of the context of the codes and putting them in the context of their particular variations---we are able to pay attention to those variations.[^49]

Take for instance, provisions regulating witness testimony. At common law, parties and interested witnesses were not permitted to testify in their own causes. Field's Code reversed this rule, expanding witness competency as widely as possible: any person "having organs of sense" could now be admitted as a witness in New York, with only the insane and very young children exempted from the rule. As the code migrated West, however, legislators added racial exclusions to Field's list. The following cluster documents how California's codifiers grafted earlier prohibitions from midwestern states into Field's Code. Many other codes then evidence a remarkable uniformity with California's text (which later changed only to add "Mongolian" to the list of races). Iowa's Code had much more minor influence, and Wyoming developed the only truly original section which made clear the link to a racist assumption of infantilization of non-white peoples.

**Figure: Cluster from Kellen**

As these clusters illustrate, uniformity can be just as instructive as variation. On the point, the most significant clusters that we investigated related to the collection of debts. These were clusters which went against the typical pattern that we observed. While many clusters exhibited regional variation as they grew more distant from the Field Code, clusters having to do with creditors' remedies were almost completely uniform across the American West and South. No single section of the code announced its preference for creditors' rights; rather, the acceleration of creditors' remedies resulted from the combination of several sections. In the original enacted Code, § 107 required a defendant to answer the complaint within twenty days, instead of at the next court session (which in some cases could have been as far as three months away); § 202 provided for default judgment as a matter of course, issued by a clerk without a judicial order if no adequate answer was received within the twenty days; §§ 128-133 abolished fictitious pleadings and required answers to state true facts verified by a defendant's oath, all so that no trial would delay the enforcement of uncontestable obligations; finally, the code abolished a traditional thirty-day waiting period between issue of judgment and commencement of execution (see First Report (New York 1848), 197). These provisions dealt with what merchants and capitalists perceived as an abuse of the common law system, where defendants in cases of debt could stretch out enforcement of debt collection for as long as two years. The Field code's summary judgment brought down the time of debt collection to a matter of weeks. Effectively, the code traded the rhythms of agriculture for the rhythms of merchant finance.[^50]

Clustering each of these sections reveals that western states along with the former Confederate states of South Carolina, North Carolina, and Florida copied each provision almost exactly. Midwestern and Upper South states that had already developed by 1850 and maintained commercial ties to Chicago and New Orleans varied the New York rules, sometimes by requiring answers only in term time, or permitting only a judge to decree default judgment rather than a clerk, in either case effectively stretching out enforcement and making a formal trial more likely. But in the Reconstruction South, and in the West over the same period, regardless of whether a jurisdiction abolished chancery or not, regardless of the racial exclusions it may have placed on witness testimony, the provisions on debt collection remained unchanged. When it came to creditors' remedies, the law of New York became the law of the land.

# Part 4

Alerted by these clustering patterns to the significance of creditors' remedies in the developing West and re-developing South, it was not difficult to find evidence of westerners and southerners commenting on these features of the code or its connections to New York capital. Twelve of the states and territories that copied the Field Code mostly closely did so during the Civil War and Reconstruction era---four states in the former Confederacy and eight jurisdictions in the Far West.[^51] As with other areas of postbellum historiography, it turns out one may learn a lot by holding the postbellum American South and American West together.

In the last decade, scholars of Reconstruction have broadened the scope of their study to include both the South and the West as two sites in one "Greater Reconstruction." These studies have illustrated the ways that military conquest, rapid industrialization, and the resettlement and education of ethnic minorities developed similarly in each region, guided by economic elites in New York and political administrators in Washington.[^52] In tracing the legal aspects of this Greater Reconstruction, scholars have focused almost entirely on constitutional rights of citizenship and civil equality or the expansion of federal power.[^53] While the 1860s and 1870s were of course a transformative period in the history of civil rights and the creation of a national state, they were also the decades in which many local legal institutions and practices were transformed not by federal power but by state codification. Naomi Lamoreaux and John Joseph Wallis have recently argued that in the creation of a modern American economy, "the federal government played *no* role in this process until the Civil War, and even then it played only a bit part." The history of the Field Code's migration helps to substantiate this claim. While Lamoreaux and Wallis focus on the development of banking, transportation, and incorporation at the state level, it was the Procedure Code that provided the structure of civil remedies that protected these state institutions. And procedure codes were, of course, the creatures of state governance.[^54]

With the universal complaints that time was too short or commissioners too democratically irresponsible, one might expect the code to have faced more competition from other workable statutes. Southern states that drew more on the civilian code tradition such as Louisiana and Alabama offered other options for states desiring to reform common law practice.[^55] And if Euro-Americans believed law had to be imported into the West, Colorado immigrants proved that the common law of Illinois was just as amenable to statutory importation as New York law.[^56] But what codifiers saw when they looked at New York, more so than Louisiana, Alabama, or Illinois, was the Empire State of commercial capital. And the putative fears, demands, and desires of a personified Capital continually wielded promises---and threats---in the debates over procedural codification. The early Mormon settlers of Utah persistently avoided the mining frenzy as well as codification. By 1870, however, the territory's governor directed the legislature's attention to the recent Code of Nevada, a code "for a people whose interests in many respects are similar to our own." Of course, standing behind the Nevada Code was "the State of New York---a State which is an empire in itself and whose commercial transactions are far greater than those of any other State in the Union." By copying its code, Utah too could be "rewarded by equal advantages."[^57] Code proponents in Colorado similarly parried arguments that a code would be a "foreign" imposition by pointing to the fact that it has been "adopted twenty-nine years ago by the Empire state of the Union" and they promised that the code of the nation's uniquely commercial empire could bring wealth in its wake.[^58]

When a Colorado legislator scoffed at the idea that capitalists could possibly care about the difference between old common law and modern code remedies, his adversaries rebuked him. "Mr. Hamill replied that he knew of one California company of capitalists who were deterred from investing in mining property here wholly on account of the practice of the courts in mining cases. If we had had this code years ago, Colorado would now have a larger amount of California capital in her mines."[^59] Codifiers argued that, in attracting capital, procedure was at least as important as the substantive rules of property and contract, because procedure secured the remedies that actually protected investments. "Men of capital and enterprise will not make investments and devote their time and energies to those works of internal improvement so necessary for the speedy development of our natural resources, and for the settlement and building up the country," Nebraska's governor reasoned, "unless ample protection is afforded them, by legal enactment, for the capital invested and labor employed." And he therefore urged swift passage of the Field Code.[^60]

Receiving innumerable letters complaining that under the code, "no one will be benefited, except perhaps some Northern Capitalists," a North Carolina commissioner undertook an anonymous defense of the new Code in the *Weekly Standard*. He encouraged the bar to accommodate themselves to change, for "the New York system ... bids fair to become national." Putatively offering an overview of the code, the articles were almost entirely about credit. "How can we create credit? By punctuality," the commissioner wrote. "And how create punctuality? by law, and by law alone. Let the law enforce punctuality; let the people of North Carolina learn that the great law of business is, that 'time is of the essence of the contract,' and incur no debt that they do not expect to meet at maturity." Under the more certain and speedy remedies of the code, "we may expect that the secret hoards of the frugal among ourselves will be offered to loan, and even that the vaults of the banks of New-York ... will be open to our industry."[^61]

As in postbellum North Carolina, establishing a flow of credit through the remedial system became a leading priority of western lawyers. While the so-called new western history has shed significant light on neglected topics of Indian dispossession and environmental management, it has often done so by leaving out of view matters of political economy, a staple of the old western history. As one work in that older tradition argued, "Debt collection was the central part of law practice for the \[western\] bar and remained a key part of private practice throughout the century."[^62] On that understanding, one western lawyer succinctly summarized the difference between the code and the common law as "whether a merchant had better try to collect a \$500 note or burn it up." Tiring of all the focus on creditors' remedies, one legislator observed that he "never knew one of these professionals who undertook to write up the beauties of the New York code, ... that he did not also break out somewhere with 'take for instance the case of an action on a promissory note,' as though the collection of notes was about all there could be any law needed for."[^63]

The creditors' remedies in the code gave the codifiers their leading argument against criticisms rooted in the ideology of popular sovereignty. "There is no doubt but the people are in favor of anything that promises to hurry up that proverbially slow and blind old female called Justice, and they will go for the old code," one Colorado newspaper announced.[^64] New York's "code practice is the best in excellence," stated another, "and when I say *best* I do not mean best for lawyers only, but best for the people---the commonwealth."[^65] If the People favored economic progress, certainty of remedy, and efficiency in proceedings, then they favored the New York Code, no matter whether they understood or cared about technical rules of pleading and joinder. Thus, in their arguments the codifiers imagined themselves the champions of popular sovereignty, for it was they who accomplished what the People actually desired.

Thus by the end of Reconstruction, New York's domestic empire of capital and creditors' remedies bore a remarkable resemblance to the international empire administered by England. Both jurisdictions, while reforming the practice of law, largely rejected codification within their own borders but encouraged it among their economic dependents. The English commissioned a full codification of law for India and colonies in Singapore, while further codifications produced by Field in New York covering civil and penal law were adopted in California and other western jurisdictions.[^66] In both places, leading arguments against codification carried a civilizational logic of empire: Advanced societies could not codify their law, for to do so would be to freeze the progress of legal science. What appeared to some to be a hopeless mass of confusion was to common law defenders the sign of true legal sophistication. Science was, after all, sophisticated.[^67] The Throop Code came in for censure, both in its first year and in following decades, precisely for trying to capture all the sophistication of the New York legal system within an unwieldy 3,300 rules.[^68] Codification, however, could help developing societies along law's frontier take a progressive leap forward. As India's chief codifier Thomas Macaulay explained, codification "cannot be well performed in an age of barbarism," but also "cannot without great difficulty be performed in an age of freedom." As India balanced between the two, however, "it is the work which especially belongs to a government like that of India---to an enlightened and paternal despotism."[^69]

In America, Macaulay's tool of enlightened despotism spread with the anxiety that capital from the nation's economic center would remain scarce without a Code of remedies that, if not in fact the law of New York, was at least prescribed by New York lawyers and their corporate clients. In the two most populous and commercially advanced western states, Texas and Illinois, the need for New York capital failed to move state legislators to adopt the code at the expense of popular sovereignty (although there were concerted efforts in both jurisdictions).[^70] Lacking the self-sufficiency of those two jurisdictions, the other states of the Greater Reconstruction in America adopted a foreign Code, but lawyers, legislators, and their supporters claimed the endorsement of popular sovereignty in doing so. Even in North Carolina, whose Democratic newspapers daily called for the repeal "this child of the carpet baggers," Republican editors proclaimed that "the movement" towards procedural codification "comes from the people, from the instinctive logic by which an unprejudiced mind grasps the advantages of the system."[^71]

# Conclusion

By framing our historical questions around "medium data" we enjoyed a useful symbiosis of traditional and digital historical methods. Our computational methods produced useful historical knowledge because they were justified by what we knew about the data from traditional historical work. We knew that code commissioners worked with "the scissors and paste-pot," as critics complained, and we knew of codes in the archives that commissions literally marked up the legislation of other states. While we think that one of the most useful things about digital history is its ability to start with large corpora and then figure out what was interesting from the past, we have shown how digital history can function by starting with particular questions rather than particular sources. We have shown how a collection of methods from computer science, including minhash/locality-sensitive hashing, affinity propagation clustering, and network analysis, along with the concept of text deformance from literary history, can be used to good effect in tracking the changes in the law, as well as any other discursive field whose texts can be readily divided into sectional units. Finally we have shown how it is possible to work on different scales, using network analysis and algorithmic close reading, and thus to get both a macroscopic picture of the law's migration, as well as a highly detailed view of the changes in the law.

The history of codification on the American periphery troubles cherished notions of American federalism. On the one hand, those suspicious of centralized power and critical of national governance since the New Deal have relied on narratives that tout the equality of sovereign states to set their own policies, foster local diversity, and "experiment" with legislative solutions to local issues.[^72] But the history of legal practice and civil remedies is one in which the localism fostered by common law practice rapidly gave way to uniform regulations transmitted by a network of New York lawyers without the slightest interference of the federal government. On the other hand, scholars who support the legitimacy of New Deal governance have recently produced an abundant literature unearthing a long tradition of "administrative" law at the state level well before the twentieth century. These accounts have focused on administrative adjudication or discretionary regulation within a narrow domain, such as customs houses.[^73] These histories have so far neglected the most widespread and significant instance of nineteenth-century administrative lawmaking in America---the spread of the Field Code through extra-legislative commissions. While these histories have sought to demonstrate that Americans from early on could be quite comfortable with administrative law and accepted it as a normal part of the constitutional order, this chapter has shown how lawmaking by commission generated significant political controversy and raised grave questions about popular sovereignty under such delegations, questions that over time were merely dodged rather than answered.

In the economically developing West and re-developing South, anxieties over the lack of capital---both among the bar as well as lay voters---joined with arguments of civilization and progress to spur many jurisdictions to copy the text of the code of New York, the Empire State of capital. The short time horizons of American lawmaking limited the options available for re-imagining or re-crafting what could become the law of remedies and legal practice. And in the economically underdeveloped parts of the country, time horizons could be short indeed. Capital might quickly pass over one region and favor another, and each month more lawyers arrived hoping to make a start in a jurisdiction where economic progress was just about to take off.

This study thus gets at the heart of lawmaking in U.S. history. Lawyers and judges, politicians and newspaper editors warred over whether codes that were drafted by commissioners and borrowed wholesale from beyond a jurisdiction's borders were actually democratic. Codifiers responded by transmuting democratic theory into support for a remedial code elected legislators had neither the time nor inclination to read. Popular support for commercial development was taken to indicate popular support for New York's civil remedies, especially the cheapened and accelerated collection of debts. In many jurisdictions the exact language of the Field Code remains on the books, and its basic provisions for civil procedure are in force throughout the United States. Without too much exaggeration we might say that our method has revealed the spine of modern American legal practice.

[^1]: Message of the Governor, in Journal of the Council for the Territory of Nevada (1862), 21.

[^2]: Nevada Council Journal (1862), 261-62. Act of Congress (1861) Organizing the Territory of Nevada, 12 U.S. Statutes at Large 209-14 (1863). 1861 Nevada Laws 314.

[^3]: Besides his work on codification, which extended to civil, penal, and even international codes of law, Field became renown for his trial advocacy. Field argued the winning side in major Reconstruction cases such as *Ex Parte Milligan*, 71 U.S. (4 Wall.) 2 (1866) (holding the trial of civilians by military commission unconstitutional), *Cummings v. Missouri*, 71 U.S. (4 Wall.) 277 (1867) (striking a loyalty oath as unconstitutional), and *United States v. Cruikshank*, 92 U.S. 542 (1875) (enforcing the Fourteenth Amendment only against "state action"). Field came under heavy censure for his representation of Gilded Age robber barons like Jay Gould, Jim Fisk, and William "Boss" Tweed, but even Tweed's chief adversary Samuel Tilden retained Field's services for the disputed election of 1876. See Henry Martyn Field, *The Life of David Dudley Field* (1898); Philip J. Bergan, "David Dudley Field: A Lawyer's Life," in *The Fields and the Law* (New York: Federal Bar Council, 1986).

[^4]: *Rocky Mountain News*, January 20, 1877.

[^5]: Roscoe Pound, "David Dudley Field: An Appraisal," and Alison Reppy, "The Field Codification Concept," in Alison Reppy, ed., *David Dudley Field: Centenary Essays* (New York University School of Law, 1949); Stephen Subrin, "David Dudley Field and the Field Code: A Historical Analysis of an Earlier Procedural Vision," 6 *Law & Hist. Rev*. 311 (1988); Robert G. Bone, "Mapping the Boundaries of a Dispute: Conceptions of Ideal Lawsuit Structure from the Field Code to the Federal Rules," 89 *Colum. L. Rev*. 1 (1989). See also the literature on procedure and codification cited below.

[^6]: When law professors such as New York's David Graham Jr. (a collaborator on the Field Code) began to be appointed to university positions, the chair for instruction in legal practice or procedure carried this designation of "practice and pleadings." Today that field is described as "civil procedure" a field that grew out of David Dudley Field's codification, as described below. The history of practice and procedure is a staple of general legal history. See Frederick Pollock and Frederick Maitland, *The History of English Law Before the Time of Edward I* (2d. ed., Cambridge, 1898), 2 vols.; Theodore Plucknett, *A Concise History of the Common Law* (5th ed. 2010); Lawrence M. Friedman, *A History of American Law* (3d. ed. 2005). Few book-length works have been dedicated to the subject, however. The exceptions are Robert Wyness Millar, *Civil Procedure of the Trial Court in Historical Perspective* (New York University School of Law 1952); Edward Purcell, *Litigation and Inequality: Federal Diversity Jurisdiction in Industrial America, 1870-1958* (Oxford 1992); John H. Langbein et al., *History of the Common Law: The Development of Anglo-American Legal Institutions* (Aspen 2009); and to some extent, William E. Nelson, *The Americanization of the Common Law: The Impact of Legal Change on Massachusetts Society, 1760-1830* (2d. ed., University of Georgia Press 1994).

[^7]: For the rise of treatises in America generally, see G. Edward White, *The Marshall Court and Cultural Change, 1815--1835* (1988) and A. W. B. Simpson, "The Rise and Fall of the Legal Treatise: Legal Principles and the Forms of Legal Literature," in *Legal Theory and Legal History: Essays on the Common Law* (London: Hambledon Press, 1987), 273--320. The most instructive treatises for New York practice in the 1840s were Oliver L. Barbour, *A Treatise on the Practice of the Court of Chancery* (1844); Alexander M. Burrill, *Treatise on the Practice of the Supreme Court of the State of New York* (1846), 2 vols.; David Graham, *A Treatise on on the Practice of the Supreme Court of the State of New York* (3d. ed. 1847); David Graham, *A Treatise on the Organization and Jurisdiction of the Courts of Law and Equity in the State of New York* (1839); Claudius L. Monell, *A Treatise on the Practice of the Supreme Court of the State of New York* (1849); Joseph W. Moulton, *The Chancery Practice of the State of New York* (1829), 2 vols.; and as a general introduction to the field, the Englishman Henry John Stephen's *Treatise on the Principles of Pleading in Civil Actions* (2d. ed. 1828).

[^8]: See Michael Lobban, *The Common Law and English Jurisprudence 1760-1850* (Clarendon Press 1991); Kunal M. Parker, *Common Law, History, and Democracy in America, 1790-1900: Legal Thought before Modernism* (Cambridge: Cambridge University Press, 2011); David M. Rabban, *Law's History: American Legal Thought and the Transatlantic Turn to History* (Cambridge 2013).

[^9]: Or, to use a term from contemporary analysis, "super statutes." William N. Eskridge and John Ferejohn, *A Republic of Statutes: The New American Constitution* (New Haven: Yale University Press, 2010). See also David Lieberman, *The Province of Legislation Determined: Legal Theory in Eighteenth-century Britain* (Cambridge 1989); Farah Peterson, Statutory Interpretation and Judicial Authority, 1776-1860 (Ph.D. dissertation, Princeton University, September 2015).

[^10]: Friedman, *A History of American Law*, 302. When the intellectual historian Perry Miller developed a reader surveying *The Legal Mind in America*, codification was its only central theme, as Miller argued it was the one intellectual topic that attracted lawyers away from their practices long enough to debate. Perry Miller, ed., *The Legal Mind in America: From Independence to the Civil War* (Anchor 1962). See also Charles M. Cook, *The American Codification Movement: A Study in Antebellum Legal Reform* (1983); Robert W. Gordon, "The American Codification Movement," 36 *Vanderbilt Law Review* 431 (1983); Maurice Eugen Lang, *Codification in the British Empire and America* (Lawbook Exchange 1924). On common theories of codification that transcended jurisdictional boundaries, see Csaba Varga, *Codification as a Socio-Historical Phenomenon* (Budapest, 2d. ed. 2011 \[1991\]); Roger Berkowitz, *The Gift of Science: Leibniz and the Modern Legal Tradition* (Fordham 2010).

[^11]: The most influential account has been Morton Horwitz's, which declares that "the desire to separate law and politics has always been a central aspiration of the American legal profession" in order to protect elite interests against popular democracy. Horwitz identifies "orthodox legal thought" and "orthodox lawyers" with the elite of the American bar who sought to shield the law from political interference, which above all meant crusading against legislation and especially codification. Morton J. Horwitz, *The Transformation of American Law, 1780-1850* (Oxford University Press, 1977), 258-59. Recent work has challenged Horwitz's account by showing how elite common law lawyers, particularly Horwitz's main target James Coolidge Carter, were actually political progressives who supported redistributive legislation such as the income tax. See, for instance, Rabban, *Law's History*, 322-77; Parker, *Common Law, History, and Democracy*, 230-41; Lewis A. Grossman, "James Coolidge Carter and Mugwump Jurisprudence," *Law & History Review* 20 (2002). These accounts follow Horwitz, however, in focusing on the few outspoken opponents of codification, rather than the elite lawyers who sponsored the procedure codes. Among the latter group could be found some of the most devout theorists of laissez faire economics in nineteenth-century America, including David Dudley Field and his brother, the Supreme Court Justice Stephen Johnson Field.

[^12]: James Fenimore Cooper, *The Ways of the Hour: A Tale* (1850), 84. For an analysis of Cooper's philosophy of law and his critique of the New York constitution, see Charles Hansford Adams, *"The Guardian of the Law": Authority and Identity in James Fenimore Cooper* (Penn State University Press (1990), 135-48. See also Marvin Meyers, *The Jacksonian Persuasion: Politics and Belief* (Stanford 1957), 57-100.

[^13]: On the politics and reforms of the New York Convention of 1846, see Charles Z. Lincoln, *The Constitutional History of New York from the Beginning of the Colonial Period to the Year 1905* (Rochester 1905), 2:10-101; Charles McCurdy, *The Anti-Rent Era in New York Law and Politics, 1839-1865* (University of North Carolina 2001); Jed Shugerman, *The People's Courts: The Rise of Judicial Elections and Judicial Power in America* (Harvard University Press 2012).

[^14]: Count Pierre François Réal to William Sampson, October 27, 1824, in Sampson's Discourse and Correspondence with Various Learned Jurists Upon the History of the Law (1826), 191; Maxwell Bloomfield, "William Sampson and the Codifiers: The Roots of American Legal Reform," 11 *American Journal of Legal History* 234 (1967); Walter J. Walsh, "William Sampson, a Republican Constitution, and the Conundrum of Orangeism on American Soil, 1824-1831," 5/7 *Radharc* 1 (2006); William Sampson, *Memoirs* (2d. ed., 1817). On French-style codifications, see Robert B. Holtman, *The Napoleonic Revolution* (Louisiana State 1981); R. H. Kilbourne, *A History of the Louisiana Civil Code* (Paul M. Herbert Law Center 1987); Brian Young, *The Politics of Codification: The Lower Canadian Civil Code of 1866* (Osgoode Society 1994); John W. Cairns, *Codification, Transplants, and History: Law Reform in Louisiana (1808) and Quebec (1866)* (Talbot 2015); Roscoe Pound, "The French Civil Code and the Spirit of Nineteenth Century Law," 35 *Boston Law Review* 79 (1955).

[^15]: Thomas J. Lockwood, "A History of Royal Commissions," 5 *Osgoode Hall Law Journal* 172 (1967); Barbara Lauriat, "'The Examination of Everything': Royal Commissions in British Legal History," 31 *Statute Law Review* 24 (2010); Joanna Innes, *Inferior Politics: Social Problems and Social Policies in Eighteenth-century Britain* (Oxford 2009).

[^16]: Second Report of the Commissioners on Practice and Pleading (New York 1849), 3-4. See also, First Report of the Commissioners on Practice and Pleadings (New York 1848), iii-iv; Third Report of the Commissioners on Practice and Pleadings (New York 1849), 3, Lillian Goldman Law Library Rare Books Collection; Final Report of the Commissioners on Practice and Pleadings, in Documents of the Assembly of the State of New York, 73d sess., vol. 2, no. 16 (New York 1850), viii.

[^17]: Report of the Commissioners on Practice and Pleadings, in New York Assembly Documents, 70th sess., vol. 2, no. 202 (1847), 3-4.

[^18]: Working under this theory, the Field commission defined the content of the modern field of *civil procedure*. While western legal systems had long distinguished between the law of persons and things on the one hand, and the law of actions (the rules of litigation) on the other, in the Anglo-American tradition, the categories remained intermixed into the nineteenth century. Whether one had a substantive legal right (to property, to marry, to an office, etc.) depended upon whether and how one would sue for a remedy to vindicate that right. Blackstone's *Commentaries* attempted to describe English law in the more European terms of persons/things/actions, and Jeremy Bentham offered a more refined terminology of "substantive" law and "procedural" or "adjective" law, but until the Field Code no Anglo-American jurist had specified with precision where the line lay between substantive and procedural law. See Lobban, *The Common Law and English Jurisprudence*, 127-131, 146-151. As Amalia Kessler notes, *Bouvier's Law Dictionary* did not even define *civil procedure* until its 1897 edition, describing the term as "rather a modern one." Amalia Kessler, "Deciding Against Conciliation: The Nineteenth-Century Rejection of a European Transplant and the Rise of a Distinctively American Ideal of Adversarial Adjudication," 10 *Theoretical Inquiries in Law* 423 (2009), 481-482; *Bouvier's Law Dictionary* (1897), 2:764. Before 1848, the term was largely restricted to French usage, and American remedial law had carried the typical designation---as it did both in Graham's treatise and professorial title---of "practice and pleadings," the name likewise given to the reform commission. When the commission designated its final draft a "Code of Civil Procedure," it marked the first American attempt to give content to this category.

[^19]: Final Report of the Commissioners, 368-378, tit. 10. See also Peter Karsten, "Enabling the Poor to Have Their Day in Court: The Sanctioning of Contingency Fee Contracts, A History to 1940," 47 *DePaul Law Review* 231 (1998); Norman Spaulding, "The Luxury of the Law: The Codification Movement and the Right to Counsel," 73 *Fordham Law Review* 983 (2004) John Leubsdorf, "Toward a History of the American Rule on Attorney Fee Recovery," 47 *Law and Contemporary Problems* 9 (1984).

[^20]: On debt collection, see Part 4 below. On homestead exemptions, see Final Report of the Commissioners, 353-354, § 839; James W. Ely, "Homestead Exemption and Southern Legal Culture," in Sally Hadden & Patricia Minter eds., *Signposts: New Directions in Southern Legal History* (University of Georgia 2013), 289-314; Paul Goodman, "The Emergence of Homestead Exemption in the United States: Accommodation and Resistance to the Market Revolution, 1840-1880," 80 *Journal of American History* 470 (1993).

[^21]: Final Report of the Commissioners, 202-203, § 506 (restricting admission as an attorney to male citizens), 110, § 251 (restricting jury service to white male citizens), 714-715, § 1708 (permitting "all persons, without exceptions" to be witnesses in civil cases).

[^22]: On the French procedure code, see C.H. van Rhee, *European Traditions in Civil Procedure* (Antwerp 2005). Bentham, the leading proponent of codification in England, argued that procedure was the one department of the law that ought to remain *un*codified. So long as the law of civil and criminal obligations and the law of property were sufficiently codified, a "natural procedure" arising from judicial discretion and flexibility would be superior to "technical" written rules. Later in his career Bentham produced the "Outlines of a Procedure Code" as a "provisional" remedy, but he insisted that a procedure code on its own could not be "invested with the form of law" without "reference to the codes of law, penal and non-penal, to which it has for its object and purpose the giving execution and effect." Although it spanned nearly 200 pages, Bentham's code favored general moral maxims over precise details, for instance: "On each occasion, have constant regard for all the several ends of justice; that is to say, minimize the sum, or the balance of evil." Jeremy Bentham, Principles of Judicial Procedure with the Outline of a Procedure Code, in John Bowring, ed., *Works of Jeremy Bentham* (2d. ed. 1843 \[1839\]), 1-189, preface and 28, ch. 7 § 1. See Lobban, *The Common Law*, 127-131.

[^23]: See Revised Statutes of the State of Indiana (1852), 2:245-320 (wills); Public Statutes of the State of Minnesota (1859), 643-647, ch. 75 (mortgages); The Code of Civil Procedure of the State of California (1880), 419-420, tit. 6, 657-659, art. 5 (corporations).

[^24]: Report of the Code Commissioners to the Eighth General Assembly of the State of Iowa (1859), 296, note to § 172 ("The right to sue, follows necessarily from the right of property.") On the significance of the Married Women's Property Acts, see, e.g., Hendrik Hartog, *Man and Wife in America: A History* (Harvard 2000), 111-113, 187-192, 290-292.

[^25]: First Report of the Commissioners, iv. For legislative debates on the Code, see "Legislative Acts and Proceedings," *Albany Evening Journal*, Mar. 31, 1848.

[^26]: Report of the Committee on the Judiciary on the Bill to Continue in Office the Commissioners on Practice and Pleadings, in New York Assembly Documents, 72d sess., vol. 3, no. 47, (New York 1849), 2, 12-15.

[^27]: Final Report of the Commissioners, 225-226, § 554. For examples of "codes" that did not alter previously enacted statutes, see, for instance, Report, Appendix to the Journals of the Senate and Assembly of the State of Tennessee (1857), 191 ("The digest presents the law substantially as it now exists in the State. I have neither felt at liberty nor deemed it advisable to innovate largely upon the established system."); 1897 New Mexico Compiled Laws 9 ("The commissioners were given no authority to revise."); 1866 Illinois Compiled Laws v ("We cannot change the text, but we can arrange and systematize the entire legislation of the state upon any given subject."); 1849 Wisconsin Revised Statutes, "Advertisement" (commission "directed the subscriber to arrange the chapters into parts and titles as he thought proper, re-arranging the order of the sections or transposing them from one chapter to another, whenever it would not alter the meaning of the law.").

[^28]: *Montana Post*, January 21, 1865.

[^29]: *The Miner's Express* (Dubuque), February 26, 1851.

[^30]: Full citations to all of the codes that we used, plus links to electronic versions at the Hathi Trust, Google Books or other sources when available are available at Kellen Funk, *American Legislation Project* (2015--): <http://kellenfunk.org/legislation>.

[^31]: We have preferred to use the term "digital history" when referring to our own work, in part because the term "digital humanities" has largely come to refer to the work of digital literary and digital media scholars, but primarily because we wish to see digital scholars make disciplinary, rather than interdisciplinary contributions. On the role of disciplines and the importance of field specific argumentation, see, Stephen Robertson, "The Differences between Digital Humanities and Digital History," {PAGE}--{PAGE}, and Cameron Blevins, "Digital History's Perpetual Future Tense," {PAGE}--{PAGE}, both in *Debates in the Digital Humanities 2016*, ed. Matthew K. Gold and Lauren F. Klein (Minneapolis: University of Minnesota Press, 2016). Nevertheless, digital text analysis in the humanities has mostly been published by literary scholars, including Stephen Ramsay, *Reading Machines: Toward an Algorithmic Criticism* (Urbana: University of Illinois Press, 2011); Franco Moretti, *Distant Reading* (New York: Verso, 2013); Matthew L. Jockers, *Macroanalysis: Digital Methods and Literary History* (Urbana: University of Illinois Press, 2013); Ted Underwood, *Why Literary Periods Mattered: Historical Contrast and the Prestige of English Studies* (Stanford: Stanford University Press, 2013); Ryan Cordell and David A. Smith, *Viral Texts: Mapping Networks of Reprinting in 19th-Century Newspapers and Magazines*, NULab for Texts Maps and Networks, Northeastern University (2012--): <http://viraltexts.org>. Related but less well regarded by humanities scholars is work in "culturomics": see Jean-Baptiste Michel et al., "Quantitative Analysis of Culture Using Millions of Digitized Books", *Science* 331, no. 6014 (2011), 176--182; doi:10.1126/science.1199644. On the uselessness of the term "big data," see Ted Underwood, "Against (Talking About) 'Big Data,'" *The Stone and the Shell*, blog post, 10 May 2013: <https://tedunderwood.com/2013/05/10/why-it-matters-that-we-dont-know-what-we-mean-by-big-data/>. For an example of curating a corpus aimed at research questions, see Ted Underwood, Boris Capitanu, Peter Organisciak, Sayan Bhattacharyya, Loretta Auvil, Colleen Fallaw, J. Stephen Downie, "Word Frequencies in English-Language Literature, 1700-1922," dataset, v0.2 (HathiTrust Research Center, 2015) doi:10.13012/J8JW8BSJ.

    Only a few scholars have turned their attention to the computer analysis of legal texts for historical purposes, including Paul Craven, "Detección automática y visualización de dominios específicos similares en documentos: análisis DWIC y su aplicación en el Proyecto Master & Servant \[Automatic Detection and Visualization of Domain-Specific Similarities in Documents: DWIC Analysis and its Application in the Master & Servant Project\]," published on CD-ROM in F. J. A Perez et al., eds., *La Historia en una nueva frontera \[History in a New Frontier\]* (Digibis: Ediciones de la Universidad de Castilla-La Mancha, 1998); Paul Craven and Douglas Hay, "Computer Applications in Comparative Historical Research: The Master & Servant Project at York University, Canada," *History and Computing* 7, no. 2 (1995); Paul Craven and W. Traves, "A General-Purpose Hierarchical Coding Engine and Its Application to Comparative Analysis of Statutes," *Literary and Linguistic Computing* 8, no. 1 (1993): 27--32, doi:10.1093/llc/8.1.27; Eric C. Nystrom and David S. Tanenhaus, "The Future of Digital Legal History: No Magic, No Silver Bullets," *American Journal of Legal History* 56, no. 1 (2016): 150--67, doi:10.1093/ajlh/njv017; Dan Cohen, Frederick Gibbs, Tim Hitchcock, Geoffrey Rockwell, et al., "Data Mining with Criminal Intent," white paper, 31 August 2011, <http://criminalintent.org>.

[^32]: {TODO: list exceptions to Google Books}.

[^33]: In each instance we downloaded an entire volume of sessions laws, statutory compilations, or single-volume codes of procedure and then cropped out irrelevant pages, marginalia and footnoted commentary, leaving only the statutory text. After several trials of various implementations of Tesseract (open source) and I.R.I.S. (proprietary) OCR programs, we determined that Nitro Pro PDF, which relies on I.R.I.S. software, offered the best OCR tool for this project. I.R.I.S. provides slightly more accurate readings of nineteenth-century typefaces than Tesseract, and Nitro Pro's implementation makes words, not characters, the fundamental unit of output. The latter feature made cropping between marginalia and the statute more reliable. We removed hyphenated line breaks and standardized spelling for common terms that evolved over the nineteenth century (e.g., *indorsement*).

[^34]: {TODO: citation on Stephen Johnson Field}

[^35]: {TODO: cite literature on West as legal void.}

[^36]: The formal definition of the Jaccard similarity for two sets, $A$ and $B$, is $$J(A, B) = \frac{ | A \cap B | }{ | A \cup B | }$$

[^37]: We implemented LSH as described in Jure Leskovec, Anand Rajaraman, and Jeff Ullman, *Mining of Massive Datasets*, 2nd ed. (Cambridge University Press, 2014), ch. 3, <http://www.mmds.org>; the algorithm was first described in Andrei Z. Broder, "On the Resemblance and Containment of Documents," in *Compression and Complexity of Sequences 1997: Proceedings*, (IEEE, 1997): 21--29, <http://gatekeeper.dec.com/ftp/pub/dec/SRC/publications/broder/positano-final-wpnums.pdf>. Other digital humanities projects, most notably *Viral Texts*, have used other means for detecting text reuse. The most prominent of these are algorithms for sequence alignment. (Our "textreuse" package for R also implements the Smith-Waterman local sequencing algorithm, derived from gene sequencing.) Yet the older and simpler LSH algorithm sufficed for our purposes because legal sources are easily divided into discrete sections which can be treated as independent documents. For other approaches, see David Bamman and Gregory Crane, "Discovering Multilingual Text Reuse in Literary Texts," white Paper, Perseus Digital Library (2009): <http://www.perseus.tufts.edu/publications/2009-Bamman.pdf>; Timothy Allen, Charles Cooney, Stéphane Douard, et al., "Plundering Philosophers: Identifying Sources of the Encyclopédie," *Journal of the Association for History and Computing* 13, no. 1 (2010): <http://hdl.handle.net/2027/spo.3310410.0013.107>; Glenn Roe, Russell Horton, and Mark Olsen, "Something Borrowed: Sequence Alignment and the Identification of Similar Passages in Large Text Collections," *Digital Studies / Le Champ numérique* 2, no. 1 (2010): <http://www.digitalstudies.org/ojs/index.php/digital_studies/article/view/190/235>; David A. Smith, Ryan Cordell, and Elizabeth Maddock Dillon, "Infectious Texts: Modeling Text Reuse in Nineteenth-Century Newspapers," in *2013 IEEE International Conference on Big Data*, 2013, 86--94, doi:10.1109/BigData.2013.6691675; David A. Smith, Ryan Cordell, Elizabeth Maddock Dillon, et al., "Detecting and Modeling Local Text Reuse," *Proceedings of IEEE/ACM Joint Conference on Digital Libraries* (IEEE Computer Society Press, 2014); Christopher Forstall, Neil Coffee, Thomas Buck, Katherine Roache, and Sarah Jacobson, "Modeling the scholars: Detecting Intertextuality through Enhanced Word-level N-gram Matching" *Digital Scholarship in the Humanities* 30, no. 4 (2015): 503--515; Douglas Ernest Duhaime, "Textual Reuse in the Eighteenth Century: Mining Eliza Haywood's Quotations," *Digital Humanities Quarterly* 10, no. 1 (2016): <http://www.digitalhumanities.org/dhq/vol/10/1/000229/000229.html>.

[^38]: We filtered this matrix based on what we knew about the process of borrowing. We removed any match below a threshold that we determined by reading pairs of matches. Because Jaccard similarity scores are symmetric, we also removed anachronistic matches. For instance, a code from 1851 obviously did not borrow from a code from 1877. Furthermore, in chains of borrowing (e.g., NY1850 → CA1851 → CA1868 → CA1872 → MT1895) the latest section might have a high similarity to all of its ancestors, but it was in fact borrowed only from the most recent parent. We therefore filtered the similarity matrix to remove matches within the same code; anachronistic matches; spurious matches beneath a certain threshold. Then if a section had multiple matches, we kept the match from the chronologically closest code, giving preference to codes from the same state, unless there was a substantially closer match from a different code.

[^39]: We have released two repositories with all the code used for this project. Lincoln Mullen, "textreuse: Detect Text Reuse and Document Similarity," R package version 0.1.3 (2015--): <https://github.com/ropensci/textreuse>, includes the LSH and other algorithms that we used for general purpose application. (This package was peer-reviewed by rOpenSci, a collective of academic developers who use the R programming language.) A second repository contains all of our code specific to the migration of the Field Code: <https://github.com/lmullen/civil-procedure-codes/>.

[^40]: This is obvious from the visualization, but also confirmed by all measures of centrality used in network analysis. A network is simply a list of edges (in our case the number of sections borrowed) between nodes (in our case, the codes). Because even our efforts at determining the best match for each section sometimes attributed a section to an incorrect code, we pruned the edges of the graph so that each code was connected to another code only if it borrowed at least fifty sections or twenty percent of its sections. On networks see {CITATION}.

    Within New York there was a definite chronological progression from the 1848, 1849, 1850, 1851, and 1853 versions of the code, but the development is not chronologically linear. The state legislature enacted the 1848, 1849, and 1851 codes, and these show stronger relational similarities. The 1850 and 1853 were David Dudley Field's ideal drafts of the code which were never enacted. They were, however, printed with wide margins, quality typesetting, and---in the 1850 draft---extensive explanatory notes, all with an eye towards other jurisdictions copying them as a model. Those two codes show stronger similarity to one another than to the enacted drafts.

[^41]: Non-Field jurisdictions occasionally exhibited a borrowing relationship within a state or across two states. In in a few unusual instances they contributed to codes which were derived from the Field Code. For example, Alabama's 1852 Code provided a few sections to Tennessee's 1858 code, and some states like Wisconsin copied, along with the Field Code, large passages of pre-code legislation from earlier in the state's history.

[^42]: spectrogram("CA1851", mutate(best\_matches, match\_code = ifelse(match\_code %in% c("NY1848", "NY1849", "NY1850"), "New York 1849/50", match\_code)), white\_list = 3)

[^43]: (The visualization here, because of difficulties in precisely matching sections, actually understates the extent of the borrowing.) **Explain** why our method errs on the side of undercounting.

[^44]: Scholars in digital literary studies often speak of the "deformance" of texts. Clustering the sections of the codes deformed them by pulling the sections out of their context within specific codes and by arranging them by topic in chronological order. Deforming the codes allowed us to see the development and spread of the law.

[^45]: Brendan J. Frey and Delbert Dueck, "Clustering by Passing Messages Between Data Points," *Science* 315 (2007): 972--976.

[^46]: Furthermore, even though the affinity propagation clus- tering algorithm did not fully converge with our peculiar dataset, it did an adequate job clustering the documents. Because there was an exemplar section for each cluster, we were able to merge clusters whose exemplars had a high Jaccard similarity score.

[^47]: We also implemented a method called the Smith Waterman algorithm, which is used for gene sequencing, but which we adapted for natural language. This is the method used by Smith et al for Viral Texts. This permits us to extract the longest matching string from two documents and so to see how the one document modified the other.

[^48]: E.g., Moretti, Distant Reading; Jockers, Macroanalysis.

[^49]: Samuels, Lisa, and Jerome McGann. "Deformance and Interpretation." *New Literary History* 30.1 (1999) : 25--56; *Stallybrass, Peter. "Against Thinking."* PMLA *122.5 (2007): 1580-1587.*

[^50]:

[^51]: Those jurisdictions were Nevada (1861), Dakota Territory---which retained the Code when split into North and South (1862), Idaho (1864), Arizona (1864), Montana (1865), Arkansas (1868), North Carolina (1868), Wyoming (1869), Florida (1870), South Carolina (1870), Utah (1870), and Colorado (1877).

[^52]: Elliott West, "Reconstructing Race," 34 *Western Historical Quarterly* 6 (2003). See also Heather Cox Richardson, *West from Appomattox: The Reconstruction of America after the Civil War* (New Haven: Yale University Press, 2007); Sven Beckert, *Monied Metropolis: New York City and the Consolidation of the American Bourgeoisie, 1850-1896* (Harvard 2003); Mark Wahlgren Summers, *The Ordeal of the Reunion: A New History of Reconstruction* (Chapel Hill: University of North Carolina Press, 2014). The major application of the Greater Reconstruction idea to legal history has been Sarah Barringer Gordon, *The Mormon Question: Polygamy and Constitutional Conflict in Nineteenth-Century America* (Chapel Hill: University of North Carolina Press, 2003).

[^53]: See, for instance, Rogers Smith, *Civic Ideals: Conflicting Visions of Citizenship in U.S. History* (Yale 1999); Meg Jacobs, William J. Novak, and Julian E. Zelizer, eds., *The Democratic Experiment: New Directions in American Political History* (Princeton 2003); Edward Purcell, *Litigation and Inequality: Federal Diversity Jurisdiction in Industrial America, 1870-1958* (Oxford 1992).

[^54]: Naomi Lamoreaux and John Joseph Wallis, "States, Not Nation: The Sources of Political and Economic Development in the Early United States," Paper Presented at the Economic History Workshop, Harvard University, March 4, 2016.

[^55]: For instance, although Tennessee in 1858 adapted nearly 225 sections of its code from Field codes, the state also incorporated nearly 50 sections of the 1852 Code of Alabama, one of the largest borrowings of southern legislation within the corpus we collected.

[^56]: Illinois was a fiercely traditional anti-code state. In later decades, lawyers jokingly called it "the Yellowstone Park of common law pleading." See Charles E. Clark, "The New Illinois Civil Practice Act," 1 *University of Chicago Law Review* 209 (1933). Nevertheless, even Illinois had substantial legislation organizing the courts and prescribing certain common law processes for obtaining civil remedies, and Colorado adapted the bulk of this legislation when it organized as a territory, a full seventeen years before its legislature considered adopting the Field Code.

[^57]: Journal of the Assembly of the Territory of Utah (1870), 15.

[^58]: "The Code Again," *Pueblo Daily Chieftain*, February 25, 1877.

[^59]: "The Legislature: The Senate Devotes Another Day to the Code," *Denver Daily Tribune*, February 17, 1877.

[^60]: Governor's Message, in Journal of the House of Assembly of the Territory of Nebraska (1857), 12.

[^61]: William A. Jenkins to William Blount Rodman, January 14, 1868, Rodman Papers. East Carolina University Library, Special Collections. Rodman's explication of the Code appeared in three sequentially numbered articles in the *Standard* on August 14, 15, and 16, 1868, under the title "The Code of Civil Procedure." Rodman disclosed his authorship in private correspondence with Barringer. See Barringer to Rodman, August 21, 1868, Rodman Papers.

[^62]: Gordon Morris Bakken, *Practicing Law in Frontier California* (Nebraska 1991), 51-54. The new western history ushered in by Patricia Nelson Limerick, *The Legacy of Conquest: The Unbroken Past of the American West* (Norton 1987) and Richard White, *"It's Your Misfortune and None of My Own": A History of the American West* (Oklahoma 1991) is now returning to issues of political economy. See Patricia Nelson Limerick, *A Ditch in Time: The City, the West, and Water* (Fulcrum 2012); Richard White, *Railroaded: The Transcontinentals and the Making of Modern America* (Norton 2011).

[^63]: "The Code," *Denver Daily Tribune*, January 10, 1877; "The Code," *Denver Daily Times*, January 27, 1877.

[^64]: "A Code of Civil Procedure," *Denver Daily Times*, January 12, 1877.

[^65]: "The Code," *Denver Daily Tribune*, January 31, 1877.

[^66]: See Gunther A. Weiss, "The Enchantment of Codification in the Common-Law World," 25 *Yale Journal of International Law* 435 (2000). For a thorough study of the ideology of codification in India, see Robert A. Yelle, *The Language of Disenchantment: Protestant Literalism and Colonial Discourse in British India* (Oxford: Oxford University Press, 2012).

[^67]: See, for instance, James C. Carter's classic defense of the common law against codification, The Ideal and the Actual in the Law, Address to the American Bar Association, August 21, 1890, at 28 ("the legislature should never attempt to perform the function of the judge, that of simply ascertaining and declaring existing customs. This is the work of experts who can qualify themselves only by the devotion of their lives.").

[^68]: See, for instance, "Note," 29 *Albany Law Journal* 141, 142 (1884); Millar, *Civil Procedure of the Trial Court*, 55-56.

[^69]: 19 *Hansard Parliamentary Debates* 531.

[^70]: Texas commissioned the preparation of a code of civil procedure in 1855, and the legislature scheduled an extra session to consider it but ultimately never passed the law. 21 *Texas Reports* (Hartley) xi (1882); *Texas State Times*, December 15, 1855. Reformers in the 1869 convention in Illinois attempted to pass a provision similar to the one in New York's 1846 constitution, which would have required the legislature to appoint a commission to revise practice and pleadings along the lines of the Field Code. See *Debates and Proceedings of the Constitutional Convention of the State of Illinois* (1870), 1496-1498.

[^71]: "The Code," *Weekly Standard* (Raleigh), May 26, 1869.

[^72]: See, for instance, the summary of the states-as-laboratories literature in Brian Galle & Joseph Leahy, "Laboratories of Democracy? Policy Innovation in Decentralized Governments," 58 *Emory Law Journal* 1333 (2009).

[^73]: See Daniel Ernst, *Tocqueville's Nightmare: The Administrative State Emerges in America, 1900-1940* (Oxford: Oxford University Press, 2014); Jerry L. Mashaw, *Creating the Administrative Constitution: The Lost One Hundred Years of American Administrative Law* (Yale Law Library 2012); Gautham Rao, *National Duties: Custom Houses and the Making of the American State* (University of Chicago 2016).