### 2018-01-08 notebooks/102-clustering.Rmd

---
title: "Cluster the documents"
author: "Lincoln Mullen"
date: "November 5, 2015"
output: html_document
---

We are going to create clusters of our documents using the affinity propagration algorithm.

Our data is bigger than anticipated when the number of section by section comparisons is stored as a non-sparse matrix. So we are going to read in the section comparisons and convert them to a sparse matrix.

```{r, message=FALSE}
library("Matrix")
library("apcluster")
library("dplyr")
library("readr")
library("textreuse")
source("R/helper.R")
load("cache/corpus-lsh.rda")
```

Filter the sections down to a meaningful value.

```{r}
scores_clustering <- scores %>% 
  filter(score >= 0.1)
scores_clustering
nrow(scores_clustering)
```

Get the sections involved:

```{r}
section_names <- lsh_subset(scores_clustering)
head(section_names, 10)
length(section_names)
```

Create a look up data frame of matrix index positions.

```{r}
lookup <- data_frame(section_names, index = 1:length(section_names))
lookup
scores_clustering <- scores_clustering %>% 
  left_join(lookup, by = c("a" = "section_names")) %>% 
  left_join(lookup, by = c("b" = "section_names")) 
scores_clustering
```

Create the sparse matrix.

```{r}
n <- length(section_names)
m <- sparseMatrix(i = scores_clustering$index.x,
                  j = scores_clustering$index.y,
                  x = scores_clustering$score,
                  dims = c(n, n), symmetric = TRUE)
colnames(m) <- section_names
rownames(m) <- section_names
```

Now use affinity propagation clustering to create clusters.

```{r}
cluster_cache <- "cache/clusters.rds"
if (!file.exists(cluster_cache)) {
  timing <- system.time(
    clu <- apcluster(s = m,
                     maxits = 100e3, convits = 10e3,
                     q = 0,
                     lam = 0.975,
                     seed = 42325, 
                     includeSim = TRUE,
                     )
  )
  require("RPushbullet")
  pbPost(title = "Clustering finished",
         body = paste0("Finished in ", timing[["elapsed"]], "."))
  saveRDS(clu, cluster_cache)
} else {
  clu <- readRDS(cluster_cache)
}
```

Get the clusters in a slightly easier format to work with.

```{r}
clusters <- clu@clusters 
names(clusters) <- names(clu@exemplars)
clusters <- lapply(clusters, names)
```

Now the problem is that some clusters have been inadequately joined together. We will compute simiarities on the examplars. If any of them have a high Jaccard similarity, we will join those clusters together.

```{r}
exemplars_corpus <- sections[names(clusters)]
exemplars_scores <- exemplars_corpus %>% 
  lsh(bands = 60) %>% 
  lsh_candidates() %>% 
  lsh_compare(exemplars_corpus, jaccard_similarity) %>% 
  arrange(desc(score))
```

Here is a box plot of those scores:

```{r}
boxplot(exemplars_scores$score)
```

What is a good minimum match? Looking at the sections, `0.2` is clearly a match, while `0.1` is clearly not. An important observation is that the exemplars for the blood quantum sections have a Jaccard similarity of `0.15`. We will use that as the threshold.

```{r}
join_threshold <- 0.15
exemplars_scores <- exemplars_scores %>% 
  filter(score >= 0.15)
```

Now a function to iterate over that data frame. If the Jaccard score is sufficiently high, combine the two clusters.

```{r}
join_clusters <- function(row) {
  exs <- c(row$a, row$b)
  minval <- which.min(extract_date(exs))
  exemplar <- exs[minval]
  duplicate <- exs[ifelse(minval == 1, 2, 1)]
  clusters[[exemplar]] <<- c(clusters[[exemplar]], clusters[[duplicate]])
  clusters[[duplicate]] <<- NULL
  dplyr::as_data_frame(row)
}

exemplars_scores %>% 
  rowwise() %>% 
  do(join_clusters(.))
```

Now convert the clusters into a data frame.

```{r}
clusters_df <- clusters %>% 
  seq_along() %>% 
  lapply(function(i) {
    exemplar <- names(clusters)[i]
    doc <- clusters[[i]]
    data_frame(exemplar, doc, cluster_id = i)
  }) %>% 
  bind_rows() %>% 
  mutate(exemplar_year = extract_date(exemplar),
         doc_date = extract_date(doc)) %>% 
  group_by(cluster_id) %>% 
  mutate(earliest_date = min(doc_date),
         earliest = doc[which.min(doc_date)]) %>% 
  select(cluster_id, exemplar, exemplar_year, earliest, earliest_date,
         doc, doc_date) %>% 
  arrange(doc_date) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  arrange(desc(n))
clusters_df
```

That data frame has the following properties. It is sorted so that the clusters with the most documents appear first. Within each cluster, documents appear chronologically. Each cluster contains three columns for documents (with another three columns corresponding to years). One column is the "examplar" document identified by the affinity propagation clustering. The next is the earliest document in the cluster. If there are multiple documents from the earliest year, then only one is listed. The third is a list of all the documents in the cluster. The column `n` is the number of documents in the cluster. All clusters regardless of length have been included.

Write that to disk:

```{r}
dir.create("out/clusters")
write_csv(clusters_df, "out/clusters/clusters.csv")
```

Some summary statistics about the borrowings.

The codes which most frequently provide exemplars, weighted by the size of the clusters for which they are exemplars.

```{r}
exemplars_summary <- clusters_df$exemplar %>% 
  extract_code_names() %>% 
  table() %>% 
  as.data.frame() %>% as_data_frame() %>% 
  rename(., code = `.`) %>% 
  arrange(desc(Freq))
write_csv(exemplars_summary, "out/clusters-exemplars-summary.csv")
exemplars_summary %>% head(20)
```

The codes which most frequently provide the earliest documents in the clusters, weighted by the size of the clusters for which they are the earliest.

```{r}
earliest_summary <- clusters_df$earliest %>% 
  extract_code_names() %>% 
  table() %>% 
  as.data.frame() %>% as_data_frame() %>% 
  rename(., code = `.`) %>% 
  arrange(desc(Freq)) 
write_csv(earliest_summary, "out/clusters-earliest-summary.csv")
earliest_summary %>% head(20)
```

Now let's write out the clusters with a certain minimum number of documents to disk.

```{r, message=FALSE}
unlink("clusters", recursive = TRUE)
dir.create("clusters", showWarnings = FALSE)

get_printable_code <- function(doc_id) {
  require("stringr")
  content <- str_wrap(content(sections[[doc_id]]))
  str_c(doc_id, content, "-----------------------------------------------\n\n", 
        sep = "\n\n")
}

get_cluster_header <- function(exemplar, earliest, n, cluster_id) {
  require("stringr")
  str_c("Exemplar: ", exemplar, "\n",
        "Earliest: ", earliest, "\n",
        "Documents in cluster: ", n, "\n",
        "Cluster ID: ", cluster_id, "\n",
        "\n",
        "-----------------------------------------------\n\n")
}

write_cluster <- function(df) {
  require("stringr")
  exemplar <- unique(df$exemplar)
  earliest <- unique(df$earliest)
  n <- unique(df$n)
  cluster_id <- unique(df$cluster_id)
  header <- get_cluster_header(exemplar, earliest, n, cluster_id)
  docs <- df$doc %>% sapply(get_printable_code) 
  text <- str_c(docs, collapse = "\n") %>% 
    str_c(header, .) %>% 
    str_split("\n") %>% 
    unlist()
  filename <- str_c("clusters/", earliest, ".txt")
  message("Writing ", filename)
  writeLines(text, filename)
  df
}

clusters_df %>% 
  filter(n >= 5) %>% 
  group_by(cluster_id) %>% 
  do(write_cluster(.))
```