### 2018-01-08 notebooks/105-most-borrowed-sections.Rmd

---
title: "Most Borrowed Sections"
author: "Kellen Funk and Lincoln Mullen"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Which sections of the Field Code were most borrowed? By borrowed we mean both directly and also several generations down the tree.

```{r, message=FALSE}
library(dplyr)
load("cache/corpus-lsh.rda")
source("R/helper.R")
source("R/spectrogram.R")
```

This is a hacky way of solving it. (The hackiness mostly comes from bending things to fit the spectrogram function.)
```{r}
borrowed_from <- function(code, matches_df = all_matches, threshold = 0.2) {
  require(dplyr)
  matches_df %>% 
    filter(match_code == code,
           borrower_year > extract_date(code),
           score >= threshold)
}

most_borrowed <- function(matches_df, matching_code, weighted = FALSE) {
  if (weighted) {
    matches_df <- matches_df %>% 
      group_by(match_section) %>% 
      summarize(score = sum(score, na.rm = TRUE)) %>% 
      mutate(match_code = "All codes",
             score = scales::rescale(score, from = c(0, max(score, na.rm = TRUE)))) %>% 
      rename(borrower_section = match_section)
  } else {
    matches_df <- matches_df %>% 
      group_by(match_section) %>% 
      summarize(n = n()) %>% 
      mutate(match_code = "All codes",
             n = ifelse(n == 1, NA, n),
             score = scales::rescale(n, from = c(0, max(n, na.rm = TRUE)))) %>% 
      rename(borrower_section = match_section)
  }

  all_sections <- best_matches %>% 
    filter(borrower_code == matching_code) %>% 
    select(borrower_section)

  all_sections %>% 
    left_join(matches_df, by = "borrower_section") %>% 
    mutate(borrower_code = matching_code)

}

borrowings_spectrogram <- function(code) {
  spectrogram(code, most_borrowed(borrowed_from(code), code, weighted = FALSE))
}
```

Most of New York 1850 is borrowed, but not the title sections as you would expect.

```{r}
borrowings_spectrogram("NY1850")
```

Some other codes:

```{r}
borrowings_spectrogram("NY1829")
borrowings_spectrogram("NY1848")
borrowings_spectrogram("MA1836")
borrowings_spectrogram("MD1855")
borrowings_spectrogram("CA1851")
```